---
title: "Correlation of PM concentrations and SARS-CoV-2 infection"
always_allow_html: true
editor_options:
  chunk_output_type: console
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Set up working environment and defaults --------------------------------------
library(envimaR)
if(Sys.info()[["nodename"]] == "PC19616"){
  source("~/plygrnd/CovidAirPolution/CovidAirPolution/src/functions/000_setup.R")
} else {
  source("~/project/cov/CovidAirPolution/src/functions/000_setup.R")  
}
compile_data = FALSE

start_date = as.POSIXct("2020-02-15")
end_date = as.POSIXct("2020-04-20")
pm = "PM10"  # PM10
Sys.setlocale("LC_TIME", "English") 

# Load Germany -----------------------------------------------------------------
if(pm == "PM2.5"){
  savefile = "germany_extended_025.RDS"
} else {
  savefile = "germany_extended_100.RDS"
}
if(compile_data){
  germany = compileDataDE(start_date = start_date, end_date = end_date,
                                  pm = pm)
  saveRDS(germany, file.path(envrmt$path_tmp, savefile))
} else {
  germany = readRDS(file.path(envrmt$path_tmp, savefile))
}

# [1] "LK Böblingen"                "LK Breisgau-Hochschwarzwald" "LK Erzgebirgskreis"          "LK Esslingen"                "LK Ludwigsburg"             
# [6] "LK Mansfeld-Südharz"         "LK Nordfriesland"            "LK Oberhavel"                "LK Schwarzwald-Baar-Kreis"   "LK Uelzen"                  
# [11] "SK Freiburg i.Breisgau"      "SK Heidelberg"               "SK Heilbronn"                "SK Pforzheim"  "LK Reutlingen"
```



# Geographical distribution of SARS-CoV2 infections in Germany
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = TRUE}
print(paste("Start date for analysis: ", start_date))
print(paste("End date for analysis:   ", end_date))
print(paste("PM values:               ", pm))

germany$de_nuts3_map$log_cases = log10(germany$de_nuts3_map$cases)
mapview(germany$de_nuts3_map, zcol = "log_cases")
```



# Countrywide temporal development of PM values and SARS-CoV2 infections in Germany
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
de_avg = germany$de_avg

de_avg_gam_time_03 = de_avg[de_avg$date >= as.POSIXct("2020-02-15") & de_avg$date <= as.POSIXct("2020-04-01"),]
set.seed(01042020)
# de_avg_gam_time = glm(new_cases ~ date + weekday , family = quasipoisson, data = de_avg_gam_time_03)
de_avg_gam_time = gam(new_cases ~ s(seq(length(date))) + weekday , family = quasipoisson, data = de_avg_gam_time_03)
de_avg_gam_time_03$predicted = predict(de_avg_gam_time, de_avg_gam_time_03, type = "response")
de_avg_gam_time_03$residuals = residuals(de_avg_gam_time, type="response")

figure_de_avg = ggplot() + 
  geom_line(data = de_avg, aes(x = date, y = pm_mean, color = "Daily mean PM10")) +
  geom_point(data = de_avg, aes(x = date, y = pm_mean, color = "Daily mean PM10")) + 
  geom_line(data = de_avg, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_point(data = de_avg, aes(x = date, y = new_cases, color = "Daily new cases")) + 
  geom_line(data = de_avg_gam_time_03, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
  geom_point(data = de_avg_gam_time_03, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
  geom_line(data = de_avg_gam_time_03, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_point(data = de_avg_gam_time_03, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype="dotted", color = "black") +
  geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype="dotted", color = "black") +
  # geom_text(aes(x = as.POSIXct("2020-03-14"), label="Start of contact restrictions \n", y=-1), colour="black", angle=90, size = 3) +
  # geom_text(aes(x = as.POSIXct("2020-03-17"), label="Start of shut down \n", y=-1), colour="black", angle=90, size = 3) +
  labs(x = "Date and day of week", y = bquote("Infections, PM10 " ~µm/m^3)) +
  theme_bw() + 
  scale_x_datetime(date_labels = "%d.%m, %a", date_breaks = "2 day", date_minor_breaks = "1 day") + 
  scale_color_manual(values=c("#b2df8a", "#1f78b4", "#a6cee3")) +
  theme(text = element_text(size = 10), axis.title = element_text(size = 10),
        legend.position = c(0.2, 0.7), axis.text.x = element_text(angle=90, hjust=0.95, vjust=0.2), 
        legend.title = element_blank(), panel.background = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank())
figure_de_avg


lims <- as.POSIXct(c(min(de_avg$date), max(de_avg$date)))

figure_de_avg_res = ggplot() + 
  geom_line(data = de_avg_gam_time_03, aes(x = date, y = residuals, color = "Daily new cases, residuals")) +
  geom_point(data = de_avg_gam_time_03, aes(x = date, y = residuals, color = "Daily new cases, residuals")) +
  geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype="dotted", color = "black") +
  geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype="dotted", color = "black") +
  geom_text(aes(x = as.POSIXct("2020-03-14"), label="Start of contact restrictions \n", y=-0.6), colour="black", angle=90, size = 2) +
  geom_text(aes(x = as.POSIXct("2020-03-17"), label="Start of shut down \n", y=-0.65), colour="black", angle=90, size = 2) +
  labs(x = "Date and day of week", y = "Residuals") +
  theme_bw() + 
  scale_x_datetime(date_labels = "%d.%m, %a", date_breaks = "2 day", date_minor_breaks = "1 day", limits = lims) +
  scale_color_manual(values=c("#d95f02")) +
  theme(text = element_text(size = 10), axis.title = element_text(size = 10),
        legend.position = "none", axis.text.x = element_text(angle=90, hjust=0.95, vjust=0.2), 
        legend.title = element_blank(), panel.background = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank())
figure_de_avg_res
```



# Dynmiac time warp clustering on countrywide average
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = TRUE}
mapview(germany$de_clstr$clstr_map, zcol = "cluster_covid")
mapview(germany$de_clstr$clstr_map, zcol = "cluster_pm")
```

```{r, eval = TRUE, echo=FALSE, warning=FALSE, message=FALSE}
de_clstr = germany$de_clstr

clstr_avg_pm_gam_time = de_clstr$clstr_avg_pm[
  de_clstr$clstr_avg_pm$date >= as.POSIXct("2020-02-15") & 
    de_clstr$clstr_avg_pm$date <= as.POSIXct("2020-04-01"),]
clstr_avg_pm_gam_time$weekday = weekdays(clstr_avg_pm_gam_time$date)
clstr_avg_pm_gam_time$predicted = NA
clstr_avg_pm_gam_time$residuals = NA


for(c in unique(clstr_avg_pm_gam_time$cluster_pm)){
  set.seed(01042020)
  gam_time = gam(new_cases ~ s(seq(length(date))) + weekday , 
                 family = quasipoisson, 
                 data = clstr_avg_pm_gam_time[clstr_avg_pm_gam_time$cluster_pm == c, ])
  clstr_avg_pm_gam_time[clstr_avg_pm_gam_time$cluster_pm == c, "predicted"] = predict(gam_time, type = "response")
  clstr_avg_pm_gam_time[clstr_avg_pm_gam_time$cluster_pm == c, "residuals"] = residuals(gam_time, type = "response")
}

clstr_avg_covid_gam_time = de_clstr$clstr_avg_covid[
  de_clstr$clstr_avg_covid$date >= as.POSIXct("2020-02-15") & 
    de_clstr$clstr_avg_covid$date <= as.POSIXct("2020-04-01"),]
clstr_avg_covid_gam_time$weekday = weekdays(clstr_avg_covid_gam_time$date)
clstr_avg_covid_gam_time$predicted = NA
clstr_avg_covid_gam_time$residuals = NA

for(c in unique(clstr_avg_covid_gam_time$cluster_covid)){
  set.seed(01042020)
  gam_time = gam(new_cases ~ s(seq(length(date))) + weekday , 
                 family = quasipoisson, 
                 data = clstr_avg_covid_gam_time[clstr_avg_covid_gam_time$cluster_covid == c, ])
  clstr_avg_covid_gam_time[clstr_avg_covid_gam_time$cluster_covid == c, "predicted"] = predict(gam_time, type = "response")
  clstr_avg_covid_gam_time[clstr_avg_covid_gam_time$cluster_covid == c, "residuals"] = residuals(gam_time, type = "response")
}


ggplot() + 
  geom_line(data = clstr_avg_pm_gam_time, aes(x = date, y = pm_mean, color = "Daily mean PM10")) + 
  geom_line(data = clstr_avg_pm_gam_time, aes(x = date, y = new_cases, color = "Daily new cases")) + 
  geom_line(data = clstr_avg_pm_gam_time, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
  geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype="dotted", color = "black") +
  geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype="dotted", color = "black") +
  labs(x = "Date and day of week", y = bquote("Infections, PM10 " ~µm/m^3)) +
  theme_bw() + 
  scale_x_datetime(date_labels = "%d.%m, %a", date_breaks = "5 day", date_minor_breaks = "1 day") + 
  scale_color_manual(values=c("#b2df8a", "#1f78b4", "#a6cee3")) +
  theme(text = element_text(size = 10), axis.title = element_text(size = 10),
        axis.text.x = element_text(angle=90, hjust=0.95, vjust=0.2), 
        legend.title = element_blank(), panel.background = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  facet_wrap(~cluster_pm, ncol =2, nrow = 6, scales = "fix")

ggplot() + 
  geom_line(data = clstr_avg_covid_gam_time, aes(x = date, y = pm_mean, color = "Daily mean PM10")) + 
  geom_line(data = clstr_avg_covid_gam_time, aes(x = date, y = new_cases, color = "Daily new cases")) + 
  geom_line(data = clstr_avg_covid_gam_time, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
  geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype="dotted", color = "black") +
  geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype="dotted", color = "black") +
  labs(x = "Date and day of week", y = bquote("Infections, PM10 " ~µm/m^3)) +
  theme_bw() + 
  scale_x_datetime(date_labels = "%d.%m, %a", date_breaks = "5 day", date_minor_breaks = "1 day") + 
  scale_color_manual(values=c("#b2df8a", "#1f78b4", "#a6cee3")) +
  theme(text = element_text(size = 10), axis.title = element_text(size = 10),
        axis.text.x = element_text(angle=90, hjust=0.95, vjust=0.2), 
        legend.title = element_blank(), panel.background = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  facet_wrap(~cluster_covid, ncol =2, nrow = 6, scales = "fix")
```



# Correlation of PM and SARS-CoV2 infections
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
de_clstr = germany$de_clstr$clstr

auxinf = lapply(unique(de_clstr$nuts3Code), function(n){
  tmp = de_clstr[de_clstr$nuts3Code == n, ]
  tmp$date_numeric = seq(nrow(tmp))
  
  tmp = data.frame(nuts3Code = tmp$nuts3Code[1],
                   date_start = tmp$date[which(tmp$cases > 0)[1]],
                   date_max = tmp$date[tail(which(tmp$new_cases == max(tmp$new_cases)), n = 1)],
                   daily_max = tmp$new_cases[tail(which(tmp$new_cases == max(tmp$new_cases)), n = 1)],
                   cases_at_max = tmp$cases[tail(which(tmp$new_cases == max(tmp$new_cases)), n = 1)],
                   cases_on_0401 = tmp$cases[tmp$date == as.POSIXct("2020-04-01")])
  
  return(tmp)
})
auxinf = do.call("rbind", auxinf)
auxinf$difftime_start_max = difftime(auxinf$date_max, auxinf$date_start, units = "days")
auxinf$first_shutdown = difftime(auxinf$date_start, as.POSIXct("2020-03-17"), units = "days")
auxinf$max_shutdown = difftime(auxinf$date_max, as.POSIXct("2020-03-17"), units = "days")

summary(as.numeric(auxinf$difftime_start_max))
summary(as.numeric(auxinf$first_shutdown))
summary(as.numeric(auxinf$max_shutdown))
summary(auxinf$daily_max)

hist(as.numeric(auxinf$difftime_start_max))
hist(as.numeric(auxinf$first_shutdown))
hist(as.numeric(auxinf$max_shutdown))

# de_model_lag = compileLaggedGLM(data = de_clstr,  
#                                 pm = "org", 
#                                 frml = "new_cases ~ date + weekday + pm_mean_lag",
#                                 nlags = 14,
#                                 subset_var = "new_cases", subset_thv = 10, individual = "start", ndays = c(0, 30))

de_model_lag = compileLaggedGAM(data = de_clstr,  
                                pm = "org", 
                                frml = "new_cases ~ s(seq(length(date))) + weekday + pm_mean_lag",
                                nlags = 14,
                                subset_var = "new_cases", subset_thv = 9, individual = "start", ndays = c(0, 30))

de_model_geo = compileLaggedGLMGeo(de_model_lag, germany$de_nuts3_map)

tstat = lapply(seq(0, -14), function(i){
  tmp = t.test(de_model_lag$t[de_model_lag$lag == i])
  data.frame(lag = i,
             t = tmp$statistic,
             p = tmp$p.value,
             estimate = tmp$estimate,
             conf_min = tmp$conf.int[1],
             conf_max = tmp$conf.int[2])
  
})
tstat = do.call("rbind", tstat)
tstat[tstat$p < 0.1,]

figure_effectsize =  ggplot(data = de_model_lag, aes(x = as.factor(lag), y = t)) + 
  geom_boxplot() + 
  theme_bw() +
  geom_hline(yintercept = c(0), linetype="dashed", color = "black") + 
  labs(x = "Time lag [days]", y = "Effect size") + 
  theme(text = element_text(size = 10), axis.title = element_text(size = 10),
        panel.background = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
figure_effectsize


ggplot(data = de_model_geo$test_geo[de_model_geo$test_geo$lag == -2,]) + 
  geom_point(aes(x = centroid_lat, y = t, color = as.factor(lag))) +
  theme_bw() + 
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "black") + 
  scale_x_continuous(breaks = seq(47, 55, 1)) + 
  scale_y_continuous(breaks = seq(-10, 20, 2)) +
  labs(title = paste("Effect size of lagged ", pm), 
       x = "Latitude [degrees]", y = "Effect size", color = "Lag of PM")
```


# Dependence of correlation of PM and SARS-CoV2 infections
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
t_matrix = reshape(de_model_lag[, c("nuts3Code", "t", "lag")], idvar = "nuts3Code", timevar = "lag", direction = "wide")
rownames(t_matrix) = t_matrix$nuts3Code

# vars = c("pm_mean", "pop_total", "pop_dens", "centroid_lat", "centroid_lon", "st_area")
# t_environment = aggregate(st_drop_geometry(de_clstr[, vars]), by = list(de_clstr$nuts3Code), FUN = mean)
# names(t_environment)[1] = "nuts3Code"
# t_environment = merge(t_environment, auxinf)

t_environment = merge(germany$de_nuts3_map, auxinf)
t_environment = st_drop_geometry(t_environment)
t_matrix = merge(t_matrix, t_environment, by = "nuts3Code")
t_matrix$pop_dens_log10 = log10(t_matrix$pop_dens)
# vars = c(vars, "date_start", "first_shutdown", "max_shutdown", "cases_at_max", "daily_max", "cases_on_0401")

# pc = prcomp(t_matrix[, 2:22])
# autoplot(pc, data = t_matrix, colour = "centroid_lat", loadings = TRUE, loadings.colour = 'blue',
#          loadings.label = TRUE)

set.seed(01042020)
rd = rda(t_matrix[, seq(which(colnames(t_matrix) == "t.0"), which(colnames(t_matrix) == "t.-14"))], 
         t_matrix[, c("pm_mean", "pop_dens_log10", "cases", "centroid_lat", "centroid_lon")])
plot(rd, type = "points", scaling = 0)

rd_scores = scores(rd)

ggplot(data = rd_scores$sites, aes(x = RDA1, y = RDA2)) +
  geom_point()

arrows(0,0,scl*rd$CCA$biplot[,1],
           scl*rd$CCA$biplot[,2],
       col="red", lwd=2,
       length = 0.2, angle=20)

scl = 5
scl.txt = 0.2

plot(rd, type = 'points', scaling = 0, xlim = c(-1.8, 1.8), ylim = c(-0.9, 0.9), arrow=TRUE)

plot(rd, xlim = c(-1.8, 1.8), ylim = c(-0.9, 0.9))

orditorp(rd, display='sp', cex=0.5, scaling=0, col='blue')
text(rd, display='cn', col='red', scaling = 1)

text(rd, display='lc', col='red')

plot(scores(rd)$sites[,2] ~
     scores(rd)$sites[,1],
     axes=T, xlab="", ylab="", cex=2,
     pch=21, col="grey", bg="grey",
     xlim=c(min(scores(rd)$sites[,1])-1, 
            max(scores(rd)$sites[,1]+2)), 
     ylim=c(min(scores(rd)$sites[,2])-1, 
            max(scores(rd)$sites[,2]+2)))

arrows(0,0,scl*rd$CCA$biplot[,1],
           scl*rd$CCA$biplot[,2],
       col="red", lwd=2,
       length = 0.2, angle=20)

text(scl.txt*(rd$CCA$biplot[,1]) +
       scl*rd$CCA$biplot[,1],
     scl.txt*(rd$CCA$biplot[,1]) +
       scl*rd$CCA$biplot[,2],vars,
     cex=1, col="red")

```


# Maps
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
# SARS-CoV2 infections on 2020-04-01
background_map = st_transform(ne_countries(scale = "large", continent = "europe", returnclass = "sf"),
                             crs = "+init=epsg:25832")
layer = st_transform(germany$de_nuts3_map, crs = "+init=epsg:25832")
map_covid_infections = ggplot(data = background_map) +
  geom_sf() + 
  geom_sf(data = layer, aes(fill = cases)) +
  scale_fill_viridis_c(trans = "log10") +
  theme(text = element_text(size = 10), axis.title = element_text(size = 10),
        legend.position="bottom") +
  labs(fill = "Infections (01.04.2020)") +
  coord_sf(xlim = c((st_bbox(layer)["xmin"] - 10000), 
                    (st_bbox(layer)["xmax"] + 10000)), 
           ylim = c((st_bbox(layer)["ymin"] - 25000),
                    (st_bbox(layer)["ymax"] + 10000)), expand = FALSE) + 
  annotation_north_arrow(location = "tl", which_north = "true",
                         height = unit(0.75, "cm"), width = unit(0.75, "cm"),
                         pad_x = unit(0.0, "in"), pad_y = unit(0.1, "in"),
                         style = north_arrow_minimal)
map_covid_infections


# Long-term PM average between 2020-02-15 and 2020-04-01
background_map = st_transform(ne_countries(scale = "large", continent = "europe", returnclass = "sf"),
                             crs = "+init=epsg:25832")
layer = st_transform(germany$de_nuts3_map, crs = "+init=epsg:25832")
map_pm_average = ggplot(data = background_map) +
  geom_sf() + 
  geom_sf(data = layer, aes(fill = pm_mean)) +
  scale_fill_viridis_c(trans = "log10") +
  theme(text = element_text(size = 10), axis.title = element_text(size = 10),
        legend.position="bottom") +
  labs(fill = "PM10 mean 15.02.2020 - 01.04.2020") +
  coord_sf(xlim = c((st_bbox(layer)["xmin"] - 10000), 
                    (st_bbox(layer)["xmax"] + 10000)), 
           ylim = c((st_bbox(layer)["ymin"] - 25000),
                    (st_bbox(layer)["ymax"] + 10000)), expand = FALSE) + 
  annotation_north_arrow(location = "tl", which_north = "true",
                         height = unit(0.75, "cm"), width = unit(0.75, "cm"),
                         pad_x = unit(0.0, "in"), pad_y = unit(0.1, "in"),
                         style = north_arrow_minimal)
map_pm_average



# RDA1 values
score_geo = cbind(t_matrix$nuts3Code, scores(rd)$sites)
colnames(score_geo)[1] = "nuts3Code"
score_geo = merge(score_geo, unique(de_clstr[, "nuts3Code"]))
st_geometry(score_geo) = score_geo$geometry

background_map = st_transform(ne_countries(scale = "large", continent = "europe", returnclass = "sf"),
                              crs = "+init=epsg:25832")
layer = st_transform(score_geo, crs = "+init=epsg:25832")
map_rda1_score = ggplot(data = background_map) +
  geom_sf() + 
  geom_sf(data = layer, aes(fill = RDA1)) +
  theme(legend.position="bottom") +
  labs(fill = "RDA1") +
  coord_sf(xlim = c((st_bbox(layer)["xmin"] - 10000), 
                    (st_bbox(layer)["xmax"] + 10000)), 
           ylim = c((st_bbox(layer)["ymin"] - 25000),
                    (st_bbox(layer)["ymax"] + 10000)), expand = FALSE) + 
  annotation_north_arrow(location = "tl", which_north = "true",
                         height = unit(0.75, "cm"), width = unit(0.75, "cm"),
                         pad_x = unit(0.0, "in"), pad_y = unit(0.1, "in"),
                         style = north_arrow_minimal)
map_rda1_score



# Significant and positive t values
p = 0.1

lag_analysis = st_set_geometry(de_model_geo$test_geo[de_model_geo$test_geo$t > 0 &
                                                de_model_geo$test_geo$p < p, ], 
                        de_model_geo$test_geo$geometry[de_model_geo$test_geo$t > 0 &
                                                de_model_geo$test_geo$p < p, ])

lag_analysis = lapply(unique(lag_analysis$nuts3Code), function(n){
  data.frame(nuts3Code = lag_analysis$nuts3Code[lag_analysis$nuts3Code == n][1],
             min_lag = max(lag_analysis$lag[lag_analysis$nuts3Code == n]),
             mean_lag = mean(lag_analysis$lag[lag_analysis$nuts3Code == n]),
             max_lag = min(lag_analysis$lag[lag_analysis$nuts3Code == n]),
             geometry = lag_analysis$geometry[lag_analysis$nuts3Code == n][1])
})
lag_analysis = do.call("rbind", lag_analysis)
lag_analysis = merge(lag_analysis, germany$de_nuts3_map[, c("nuts3Code", "pm_mean", "cases", "pop_dens")])
hist(lag_analysis$min_lag)
hist(lag_analysis$mean_lag)
hist(lag_analysis$max_lag)

non_significant_regions = germany$de_nuts3_map[!(germany$de_nuts3_map$nuts3Code %in% lag_analysis$nuts3Code),]

summary(germany$de_nuts3_map[,c("nuts3Code", "pm_mean", "cases", "pop_dens")])
summary(lag_analysis)
summary(non_significant_regions[,c("nuts3Code", "pm_mean", "cases", "pop_dens")])

quantile(non_significant_regions$pop_dens, probs = seq(0, 1, 0.1))
quantile(lag_analysis$pop_dens, probs = seq(0, 1, 0.1))

wilcox.test(non_significant_regions$cases, lag_analysis$cases)
wilcox.test(non_significant_regions$pm_mean, lag_analysis$pm_mean)
wilcox.test(non_significant_regions$pop_dens, lag_analysis$pop_dens)

t.test(non_significant_regions$cases, lag_analysis$cases)
t.test(non_significant_regions$pm_mean, lag_analysis$pm_mean)
t.test(non_significant_regions$pop_dens, lag_analysis$pop_dens)

layer1 = st_set_geometry(lag_analysis, lag_analysis$geometry)
layer1 = st_transform(layer1, crs = "+init=epsg:25832")

background_map = st_transform(ne_countries(scale = "large", continent = "europe", returnclass = "sf"),
                              crs = "+init=epsg:25832")
layer = st_transform(germany$de_nuts3_map, crs = "+init=epsg:25832")

map_t_sig = ggplot(data = background_map) +
  geom_sf() + 
  geom_sf(data = layer) +
  geom_sf(data = layer1, aes(fill = mean_lag)) +
  scale_fill_viridis_c() +
  theme(legend.position="bottom") +
  labs(fill = "Mean lag of positive effect size") +
  coord_sf(xlim = c((st_bbox(layer)["xmin"] - 10000), 
                    (st_bbox(layer)["xmax"] + 10000)), 
           ylim = c((st_bbox(layer)["ymin"] - 25000),
                    (st_bbox(layer)["ymax"] + 10000)), expand = FALSE) + 
  annotation_north_arrow(location = "tl", which_north = "true",
                         height = unit(0.5, "cm"), width = unit(0.5, "cm"),
                         pad_x = unit(0.0, "in"), pad_y = unit(0.1, "in"),
                         style = north_arrow_minimal)
map_t_sig
```


# Figures for publication
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}


figure_effectsize
plot(rd)
map_covid_infections
map_pm_average


figure_1_1 = plot_grid(figure_de_avg, figure_de_avg_res, 
                       labels = c("a", "b"), label_size = 10, ncol = 1)
figure_1_2 = plot_grid(map_covid_infections, map_pm_average, 
                       labels = c("c", "d"), label_size = 10, ncol = 2)
figure_1 = plot_grid(figure_1_1, figure_1_2, 
                     labels = NULL, ncol = 1)

ggsave(file.path(envrmt$path_figures, "figure_1.png"), plot = figure_1, 
       width = 216, height = 279, units = "mm", dpi = 300)



pdf(NULL, width=5, height=10)
dev.control(displaylist="enable")
plot(rd, type = "points", scaling = 0)
figure_rda = recordPlot()
invisible(dev.off())

figure_2_1 = plot_grid(figure_effectsize, figure_rda, 
                       labels = c("a", "b"), label_size = 10, 
                       rel_widths = c(1,2))

figure_2_2 = plot_grid(map_rda1_score, map_t_sig, 
                       labels = c("c", "d"), label_size = 10, ncol = 2)

figure_2 = plot_grid(figure_2_1, figure_2_2, 
                     labels = NULL, ncol = 1)


ggsave(file.path(envrmt$path_figures, "figure_2.png"), plot = figure_2, 
       width = 216, height = 279, units = "mm", dpi = 300)

```


```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
de_clstr = germany$de_clstr$clstr
# pdata = compilePopulation()
# de_clstr = merge(de_clstr, pdata, by = "nuts3Code")
# 
# de_clstr = merge(de_clstr, st_set_geometry(germany$de_nuts3_map[, c("nuts3Code", "nuts3Name", "centroid_lat", "centroid_lon", "nuts3Area")], NULL), by.x = "nuts3Code", by.y = "nuts3Code")
# 
# de_clstr$nuts3Area = as.numeric(de_clstr$nuts3Area)
# de_clstr$cases_per_area = de_clstr$cases / de_clstr$nuts3Area
# de_clstr$cases_per_pop = de_clstr$cases / de_clstr$pop_total
# de_clstr$cases_per_pop_dens = de_clstr$cases / de_clstr$pop_dens
# de_clstr$new_cases_per_area = de_clstr$new_cases / de_clstr$nuts3Area
# de_clstr$new_cases_per_pop = de_clstr$new_cases / de_clstr$pop_total
# de_clstr$new_cases_per_pop_dens = de_clstr$new_cases / de_clstr$pop_dens
# de_clstr$pm_mean_dt = c(0, diff(de_clstr$pm_mean))
# de_clstr$weekday_cn = as.numeric(de_clstr$weekday_c)
# de_clstr$days_before_shutdown = difftime(de_clstr$date, as.POSIXct("2020-03-17"), units = "days")


info = lapply(unique(de_clstr$nuts3Code), function(n){
  tmp = de_clstr[de_clstr$nuts3Code == n, ]
  tmp$date_numeric = seq(nrow(tmp))
  
  tmp = data.frame(nuts = tmp$nuts3Code[1],
                   date_start = tmp$date[which(tmp$cases > 0)[1]],
                   date_max = tmp$date[tail(which(tmp$new_cases == max(tmp$new_cases)), n = 1)],
                   daily_max = tmp$new_cases[tail(which(tmp$new_cases == max(tmp$new_cases)), n = 1)],
                   cases_at_max = tmp$cases[tail(which(tmp$new_cases == max(tmp$new_cases)), n = 1)])
  
  # set.seed(01042020)
  # # glm_mod = glm(cases ~ date_numeric + I(date_numeric^2),
  # #               family = gaussian, 
  # #               data = tmp[tmp$weekday_c == "W" & tmp$date >= date_start,])
  # glm_mod <- zeroinfl(new_cases ~ date_numeric + weekday_c + pm_mean | days_before_shutdown, data = tmp)
  # tmp$cases_pred = 0
  # tmp$cases_pred[tmp$date >= date_start] = predict(glm_mod, tmp[tmp$date >= date_start,])
  # tmp$cases_pred[tmp$cases_pred < 0] = 0
  # plot(tmp$cases_pred)
  # points(tmp$new_cases, col = "red")
  
  return(tmp)
})
info = do.call("rbind", info)
info$difftime = difftime(info$date_max, info$date_start, units = "days")
ggplot(data = info, aes(x = difftime)) + geom_histogram(binwidth = 5)
summary(as.numeric(info$difftime[info$cases_at_max > 100]))
summary(info$cases_at_max)
summary(info$daily_max)
summary(info$cases_at_max[info$difftime>=30])
summary(info$cases_at_max)
summary(as.numeric(info$difftime))

head(de_clstr)

ggplot(data = de_clstr[de_clstr$weekday_c == "M",], aes(x = date, y = new_cases, color = as.factor(nuts3Code))) + geom_line() +  theme(legend.position = "none")

ndays = c(-14, 35)
ndays = c(-14, 20)
ndays = c(0, 30)

de_model_lag = compileLaggedGLM(data = de_clstr[de_clstr$nuts3Code %in% info$nuts[info$difftime>=30 & info$difftime<=99],],  
                                pm = "org", 
                                frml = "new_cases ~ date + weekday_c + pm_mean_lag",
                                nlags = 20,
                                subset_var = "new_cases", subset_thv = 1, individual = "start", ndays = c(0, 35))

de_model_lag = compileLaggedGAM(data = de_clstr,  
                                pm = "org", 
                                frml = "new_cases ~ seq(length(date)) + weekday_c + pm_mean_lag",
                                nlags = 20,
                                subset_var = "new_cases", subset_thv = 1, individual = "start", ndays = c(0, 35))



de_model_lag = compileLaggedGLM(data = de_clstr,  
                                pm = "org", 
                                frml = "new_cases ~ date + weekday_c + pm_mean_lag",
                                nlags = 20,
                                subset_var = "new_cases", subset_thv = 1, individual = "start", ndays = c(0, 35))



de_model_geo = compileLaggedGLMGeo(de_model_lag, germany$de_nuts3_map)

t.test(de_model_lag$t[de_model_lag$lag == -3])

# de_model_lag = de_model_lag_org
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
ggplot(data = de_model_lag[abs(de_model_lag$t)<5,], aes(x = as.factor(lag), y = t)) + 
  geom_boxplot() + 
  stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="red", fill="red") +
  theme_bw() +
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "red") + 
  labs(title = paste("Dependence of individual COVID-19 time series and lagged ", pm, " concentrations"),x = "Time lag PM [days]", y = "t value of the partial coefficient")
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
ggplot(data = de_model_lag[abs(de_model_lag$t)<5,], aes(x = as.factor(lag), y = t, fill = cluster_pm)) + 
  geom_boxplot() + 
  theme_bw() + 
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "red") + 
  labs(title = paste("Dependence of individual COVID-19 time series and lagged ", pm, " concentrations by PM clusters"),x = "Time lag PM [days]", y = "t value of the partial coefficient")
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
ggplot(data = de_model_geo$test_geo_lag) + 
  geom_line(aes(x = lag, y = rsq, color = "R squared")) +
  geom_point(aes(x = lag, y = rsq, color = "R squared")) +
  theme_bw() + 
  scale_x_continuous(breaks = seq(-15, 0, 1), minor_breaks = seq(-15, 0, 1)) + 
  labs(title = paste("R squared values of lagged ", pm, " concentrations based on latitude, longitude and area",x = "Time lag PM [days]"), y = "R squared", color = "R squared")

ggplot(data = de_model_geo$test_geo_lag) + 
  geom_line(aes(x = lag, y = centroid_lat_t, color = "latitude")) +
  geom_point(aes(x = lag, y = centroid_lat_t, color = "latitude")) +
  geom_line(aes(x = lag, y = centroid_lon_t, color = "longitude")) +
  geom_point(aes(x = lag, y = centroid_lon_t, color = "longitude")) +
  geom_line(aes(x = lag, y = area_t, color = "log10(area)")) +
  geom_point(aes(x = lag, y = area_t, color = "log10(area)")) +
  theme_bw() + 
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "black") + 
  scale_x_continuous(breaks = seq(-15, 0, 1), minor_breaks = seq(-15, 0, 1)) + 
  scale_y_continuous(breaks = seq(-10, 20, 2)) +
  labs(title = paste("t values of lagged ", pm, " concentrations for latitude, longitude and area"),x = "Time lag PM [days]", y = "t value of the partial coefficient", color = "Partial coefficient")

ggplot(data = de_model_geo$test_geo[de_model_geo$test_geo$lag == -9,]) + 
  geom_point(aes(x = centroid_lat, y = t, color = as.factor(lag))) +
  theme_bw() + 
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "black") + 
  scale_x_continuous(breaks = seq(47, 55, 1)) + 
  scale_y_continuous(breaks = seq(-10, 20, 2)) +
  labs(title = paste("t values of lagged ", pm, " concentrations for latitude"),x = "Latitude [degrees]", y = "t value of the partial coefficient for PM (glm model)", color = "Lag of PM")

ggplot(data = de_model_geo$test_geo[de_model_geo$test_geo$lag >= -7 & de_model_geo$test_geo$lag <= -4,]) + 
  geom_point(aes(x = centroid_lat, y = t, color = as.factor(lag))) +
  theme_bw() + 
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "black") + 
  scale_x_continuous(breaks = seq(47, 55, 1)) + 
  scale_y_continuous(breaks = seq(-10, 20, 2)) +
  labs(title = paste("t values of lagged ", pm, " concentrations for latitude"),x = "Latitude [degrees]", y = "t value of the partial coefficient for PM (glm model)", color = "Lag of PM")
```

Hier mal ein Versuch, die "Italien-Grafik" aus dem  etwas nachzubauen. Die Signifikanz ist bei PM10 für längere Zeitintervalle gegeben, für PM2.5 praktisch nicht (Setti, L. et al. The Potential role of Particulate Matter in the Spreading of COVID-19 in Northern Italy: First Evidence-based Research Hypotheses. medRxiv, https://www.medrxiv.org/content/10.1101/2020.04.11.20061713v1 (2020).)

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
# 10, 12
test = germany$de_clstr$clstr[germany$de_clstr$clstr$date < as.POSIXct("2020-03-17"),] %>%
  group_by(nuts3Code) %>%
  summarize(n_days = n(),
            pm_gt_10 = sum(pm_mean > 10),
            pm_gt_10_rel = pm_gt_10 / n_days)
test_merge = merge(test, germany$de_clstr$clstr[germany$de_clstr$clstr$date == as.POSIXct("2020-04-01"),])
test_merge = merge(test_merge, st_set_geometry(germany$de_nuts3_map, NULL)[, c("nuts3Code", "area")])
test_merge$ratio = as.numeric(test_merge$cases/test_merge$area)

ggplot(data = test_merge, aes(x = pm_gt_10, y = ratio)) + 
  geom_point() +
  geom_smooth(method = "glm") +
  theme_bw() + 
  scale_y_log10() +
  labs(title = paste("COVID-19 infections per province area (01.04.) and days exceeding" , pm, " > 10 from 15.2. to 17.03."))
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = TRUE}
model = glm(test_merge$ratio ~ test_merge$pm_gt_10, family = quasipoisson)
summary(model)
```



```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
# Wavelet coherence analysis

## Wavelet coherence analysis on countrywide average
de_avg = germany$de_avg
wc = compileWC(de_avg)

wc.image(wc, which.image = "wc",
         n.levels = 250, color.key = "interval",
         siglvl.contour = 0.1, siglvl.arrow = 0.1,
         legend.params = list(lab = "cross-wavelet power levels"), show.date = TRUE,
         spec.period.axis = list(at = seq(15)),
         main = paste("Wavelet coherence Germany, ", pm, " x detrended daily COVID-19 cases"))

# wc.sel.phases(wc, sel.period = 4, siglvl = 1, show.date = TRUE)
```



```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
## Wavelet coherence analysis on cluster averages
de_clstr = germany$de_clstr
for(c in unique(de_clstr$clstr_avg_detr$cluster_pm)){
  tmp = de_clstr$clstr_avg_detr[de_clstr$clstr_avg_detr$cluster_pm == c,]
  wc = compileWC(tmp)
  
  wc.image(wc, which.image = "wc",
           n.levels = 250, color.key = "interval",
           siglvl.contour = 0.1, siglvl.arrow = 0.1,
           legend.params = list(lab = "cross-wavelet power levels"), show.date = TRUE,
           spec.period.axis = list(at = seq(15)),
           main = paste("Germany, ", pm, " x detrended daily COVID-19 cases, cluster ", c))
  
  # wc.sel.phases(wc, sel.period = 4, siglvl = 1, show.date = TRUE)
}
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
de_avg_glm_time_03 = de_avg[de_avg$date >= as.POSIXct("2020-03-01") & de_avg$date <= as.POSIXct("2020-04-20"),]
set.seed(01042020)
de_avg_glm_time = gam(new_cases ~ s(seq(length(date))) + weekday, family = quasipoisson, data = de_avg_glm_time_03)
de_avg_glm_time = glm(new_cases ~ seq(length(date)) + weekday, family = quasipoisson, data = de_avg_glm_time_03)
de_avg_glm_time_03$predicted = predict(de_avg_glm_time, de_avg_glm_time_03, type = "response")
de_avg_glm_time_03$residuals = residuals(de_avg_glm_time, type="response")

figure_01 = ggplot() + 
  geom_line(data = de_avg, aes(x = date, y = pm_mean, color = "Daily mean PM10")) +
  geom_point(data = de_avg, aes(x = date, y = pm_mean, color = "Daily mean PM10")) + 
  geom_line(data = de_avg, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_point(data = de_avg, aes(x = date, y = new_cases, color = "Daily new cases")) + 
  geom_line(data = de_avg_glm_time_03, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
  geom_point(data = de_avg_glm_time_03, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
  geom_line(data = de_avg_glm_time_03, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_point(data = de_avg_glm_time_03, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype="dotted", color = "black") +
  geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype="dotted", color = "black") +
  # geom_text(aes(x = as.POSIXct("2020-03-14"), label="Start of contact restrictions \n", y=-1), colour="black", angle=90, size = 3) +
  # geom_text(aes(x = as.POSIXct("2020-03-17"), label="Start of shut down \n", y=-1), colour="black", angle=90, size = 3) +
  labs(x = "Date and day of week", y = expression("Infections, PM10 ~µm/m^{3}")) +
  theme_bw() + 
  scale_x_datetime(date_labels = "%d.%m, %a", date_breaks = "2 day", date_minor_breaks = "1 day") + 
  scale_color_manual(values=c("#b2df8a", "#1f78b4", "#a6cee3")) +
  theme(text = element_text(size = 10), axis.title = element_text(size = 10),
        legend.position = c(0.35, 0.75), axis.text.x = element_text(angle=90, hjust=0.95, vjust=0.2), 
        legend.title = element_blank(), panel.background = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank())
figure_01

```