---
title: "Correlation of PM concentrations and SARS-CoV-2 infection"
always_allow_html: true
editor_options:
  chunk_output_type: console
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Set up working environment and defaults --------------------------------------
library(envimaR)
if(Sys.info()[["nodename"]] == "PC19616"){
  source("~/plygrnd/CovidAirPolution/CovidAirPolution/src/functions/000_setup.R")
} else {
  source("~/project/cov/CovidAirPolution/src/functions/000_setup.R")  
}
compile_data = FALSE

start_date = as.POSIXct("2020-01-15")
end_date = as.POSIXct("2020-04-20")
pm = "PM10"  # PM10

# Load Germany -----------------------------------------------------------------
if(pm == "PM2.5"){
  savefile = "germany_extended_025.RDS"
} else {
  savefile = "germany_extended_100.RDS"
}
if(compile_data){
  germany = compileDataDE(start_date = start_date, end_date = end_date,
                                  pm = pm)
  saveRDS(germany, file.path(envrmt$path_tmp, savefile))
} else {
  germany = readRDS(file.path(envrmt$path_tmp, savefile))
}

# [1] "LK Böblingen"                "LK Breisgau-Hochschwarzwald" "LK Erzgebirgskreis"          "LK Esslingen"                "LK Ludwigsburg"             
# [6] "LK Mansfeld-Südharz"         "LK Nordfriesland"            "LK Oberhavel"                "LK Schwarzwald-Baar-Kreis"   "LK Uelzen"                  
# [11] "SK Freiburg i.Breisgau"      "SK Heidelberg"               "SK Heilbronn"                "SK Pforzheim"  
```

# Geographical distribution of covid cases

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = TRUE}
print(paste("Start date for analysis: ", start_date))
print(paste("End date for analysis:   ", end_date))
print(paste("PM values:               ", pm))

germany$de_nuts3_map$log_cases = log10(germany$de_nuts3_map$cases)
mapview(germany$de_nuts3_map, zcol = "log_cases")
```


# Countrywide temporal development of PM values and COVID-19 cases in Germany

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
de_avg = germany$de_avg
ggplot() + 
  geom_line(data = de_avg, aes(x = date, y = pm_mean, color = "PM")) + 
  geom_line(data = de_avg, aes(x = date, y = new_cases, color = "New cases")) + 
  geom_line(data = de_avg, aes(x = date, y = new_cases_detr, color = "New cases detrended")) + 
  geom_line(data = de_avg, aes(x = date, y = new_cases_smooth, color = "New cases smoothed")) + 
  geom_line(data = de_avg, aes(x = date, y = new_cases_smooth_detr, color = "New cases smoothed detrended")) + 
  labs(title = paste(pm, " x daily new Covid-19 cases"), x = "Date", y = "Value") +
  geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype="dotted", color = "black") +
  geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype="dotted", color = "black") +
  theme_bw() + 
  scale_x_datetime(date_labels = "%d.%m - %a", date_breaks = "1 day") + 
  # scale_color_manual(values=c("red", "blue", "darkgreen")) +
  theme(axis.text.x = element_text(angle=90)) + 
  theme(legend.title = element_blank())
```


```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
de_avg = germany$de_avg
ggplot() + 
  geom_line(data = de_avg, aes(x = date, y = pm_mean, color = "PM")) + 
  geom_line(data = de_avg, aes(x = date, y = new_cases_smooth_detr, color = "New cases (smoothed, detrended)")) +
  geom_line(data = de_avg, aes(x = date, y = new_cases_smooth, color = "New cases (smoothed)")) + 
  labs(title = paste(pm, " x daily new Covid-19 cases (smoothed, detrended)"), x = "Date", y = "Value") +
  theme_bw() + 
  scale_x_datetime(date_labels = "%d.%m - %a", date_breaks = "1 day") + 
  scale_color_manual(values=c("red", "darkgreen", "blue")) +
  theme(axis.text.x = element_text(angle=90)) + 
  theme(legend.title = element_blank())
```


# Dynmiac time warp clustering on countrywide average

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = TRUE}
mapview(germany$de_clstr$clstr_map, zcol = "cluster_covid")
mapview(germany$de_clstr$clstr_map, zcol = "cluster_pm")
```

```{r, eval = TRUE, echo=FALSE, warning=FALSE, message=FALSE}
de_clstr = germany$de_clstr
ggplot() + 
  geom_line(data = de_clstr$clstr_avg_detr, aes(x = date, y = pm_mean, color = "PM")) + 
  geom_line(data = de_clstr$clstr_avg_detr, aes(x = date, y = new_cases, color = "New cases")) + 
  geom_line(data = de_clstr$clstr_avg_detr, aes(x = date, y = new_cases_smooth_detr*5, color = "New cases (smoothed, detrended)")) +
  geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype="dotted", color = "black") +
  geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype="dotted", color = "black") +
  theme_bw() + 
  labs(title = paste(pm, " x daily new Covid-19 cases (smoothed, detrended)"), x = "Date", y = "Value") +
  scale_x_datetime(date_labels = "%d.%m - %a", date_breaks = "1 day") + 
  scale_color_manual(values=c("red", "darkgreen", "blue")) +
  theme(axis.text.x = element_text(angle=90)) + 
  theme(legend.title = element_blank()) +
  facet_wrap(~cluster_pm, ncol =2, nrow = 2, scales = "fix")
```


# Explanatory potential of PM for COVID-19 cases

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
de_clstr = germany$de_clstr$clstr
pdata = compilePopulation()
de_clstr = merge(de_clstr, pdata, by = "nuts3Code")

de_clstr = merge(de_clstr, st_set_geometry(germany$de_nuts3_map[, c("nuts3Code", "nuts3Name", "centroid_lat", "centroid_lon", "nuts3Area")], NULL), by.x = "nuts3Code", by.y = "nuts3Code")

de_clstr$nuts3Area = as.numeric(de_clstr$nuts3Area)
de_clstr$cases_per_area = de_clstr$cases / de_clstr$nuts3Area
de_clstr$cases_per_pop = de_clstr$cases / de_clstr$pop_total
de_clstr$cases_per_pop_dens = de_clstr$cases / de_clstr$pop_dens
de_clstr$new_cases_per_area = de_clstr$new_cases / de_clstr$nuts3Area
de_clstr$new_cases_per_pop = de_clstr$new_cases / de_clstr$pop_total
de_clstr$new_cases_per_pop_dens = de_clstr$new_cases / de_clstr$pop_dens
de_clstr$pm_mean_dt = c(0, diff(de_clstr$pm_mean))

ggplot(de_clstr[de_clstr$nuts3Code == 12064, ], aes(x = date, y = cases, colour = as.factor(nuts3Code))) + geom_line()


de_model_lag = compileLaggedGLM(data = de_clstr,  
                                target = "new_cases", nlags = 35,
                                min_cases = 1, individual = TRUE, ndays = c(-6, 24))

de_model_geo = compileLaggedGLMGeo(de_model_lag, germany$de_nuts3_map)

t.test(de_model_geo$test_geo[de_model_geo$test_geo$lag == -6 & de_model_geo$test_geo$centroid_lat >= 30, "t"])
t.test(de_model_lag$t[de_model_lag$lag == -3])

# de_model_lag = de_model_lag_org
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
ggplot(data = de_model_lag[abs(de_model_lag$t)<5,], aes(x = as.factor(lag), y = t)) + 
  geom_boxplot() + 
  stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="red", fill="red") +
  theme_bw() +
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "red") + 
  labs(title = paste("Dependence of individual COVID-19 time series and lagged ", pm, " concentrations"),x = "Time lag PM [days]", y = "t value of the partial coefficient")
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
ggplot(data = de_model_lag[abs(de_model_lag$t)<5,], aes(x = as.factor(lag), y = t, fill = cluster_pm)) + 
  geom_boxplot() + 
  theme_bw() + 
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "red") + 
  labs(title = paste("Dependence of individual COVID-19 time series and lagged ", pm, " concentrations by PM clusters"),x = "Time lag PM [days]", y = "t value of the partial coefficient")
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
ggplot(data = de_model_geo$test_geo_lag) + 
  geom_line(aes(x = lag, y = rsq, color = "R squared")) +
  geom_point(aes(x = lag, y = rsq, color = "R squared")) +
  theme_bw() + 
  scale_x_continuous(breaks = seq(-15, 0, 1), minor_breaks = seq(-15, 0, 1)) + 
  labs(title = paste("R squared values of lagged ", pm, " concentrations based on latitude, longitude and area",x = "Time lag PM [days]"), y = "R squared", color = "R squared")

ggplot(data = de_model_geo$test_geo_lag) + 
  geom_line(aes(x = lag, y = centroid_lat_t, color = "latitude")) +
  geom_point(aes(x = lag, y = centroid_lat_t, color = "latitude")) +
  geom_line(aes(x = lag, y = centroid_lon_t, color = "longitude")) +
  geom_point(aes(x = lag, y = centroid_lon_t, color = "longitude")) +
  geom_line(aes(x = lag, y = area_t, color = "log10(area)")) +
  geom_point(aes(x = lag, y = area_t, color = "log10(area)")) +
  theme_bw() + 
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "black") + 
  scale_x_continuous(breaks = seq(-15, 0, 1), minor_breaks = seq(-15, 0, 1)) + 
  scale_y_continuous(breaks = seq(-10, 20, 2)) +
  labs(title = paste("t values of lagged ", pm, " concentrations for latitude, longitude and area"),x = "Time lag PM [days]", y = "t value of the partial coefficient", color = "Partial coefficient")

ggplot(data = de_model_geo$test_geo[de_model_geo$test_geo$lag == -9,]) + 
  geom_point(aes(x = centroid_lat, y = t, color = as.factor(lag))) +
  theme_bw() + 
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "black") + 
  scale_x_continuous(breaks = seq(47, 55, 1)) + 
  scale_y_continuous(breaks = seq(-10, 20, 2)) +
  labs(title = paste("t values of lagged ", pm, " concentrations for latitude"),x = "Latitude [degrees]", y = "t value of the partial coefficient for PM (glm model)", color = "Lag of PM")

ggplot(data = de_model_geo$test_geo[de_model_geo$test_geo$lag >= -7 & de_model_geo$test_geo$lag <= -4,]) + 
  geom_point(aes(x = centroid_lat, y = t, color = as.factor(lag))) +
  theme_bw() + 
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "black") + 
  scale_x_continuous(breaks = seq(47, 55, 1)) + 
  scale_y_continuous(breaks = seq(-10, 20, 2)) +
  labs(title = paste("t values of lagged ", pm, " concentrations for latitude"),x = "Latitude [degrees]", y = "t value of the partial coefficient for PM (glm model)", color = "Lag of PM")
```

Hier mal ein Versuch, die "Italien-Grafik" aus dem  etwas nachzubauen. Die Signifikanz ist bei PM10 für längere Zeitintervalle gegeben, für PM2.5 praktisch nicht (Setti, L. et al. The Potential role of Particulate Matter in the Spreading of COVID-19 in Northern Italy: First Evidence-based Research Hypotheses. medRxiv, https://www.medrxiv.org/content/10.1101/2020.04.11.20061713v1 (2020).)

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
# 10, 12
test = germany$de_clstr$clstr[germany$de_clstr$clstr$date < as.POSIXct("2020-03-17"),] %>%
  group_by(nuts3Code) %>%
  summarize(n_days = n(),
            pm_gt_10 = sum(pm_mean > 10),
            pm_gt_10_rel = pm_gt_10 / n_days)
test_merge = merge(test, germany$de_clstr$clstr[germany$de_clstr$clstr$date == as.POSIXct("2020-04-01"),])
test_merge = merge(test_merge, st_set_geometry(germany$de_nuts3_map, NULL)[, c("nuts3Code", "area")])
test_merge$ratio = as.numeric(test_merge$cases/test_merge$area)

ggplot(data = test_merge, aes(x = pm_gt_10, y = ratio)) + 
  geom_point() +
  geom_smooth(method = "glm") +
  theme_bw() + 
  scale_y_log10() +
  labs(title = paste("COVID-19 infections per province area (01.04.) and days exceeding" , pm, " > 10 from 15.2. to 17.03."))
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = TRUE}
model = glm(test_merge$ratio ~ test_merge$pm_gt_10, family = quasipoisson)
summary(model)
```



```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
# Wavelet coherence analysis

## Wavelet coherence analysis on countrywide average
de_avg = germany$de_avg
wc = compileWC(de_avg)

wc.image(wc, which.image = "wc",
         n.levels = 250, color.key = "interval",
         siglvl.contour = 0.1, siglvl.arrow = 0.1,
         legend.params = list(lab = "cross-wavelet power levels"), show.date = TRUE,
         spec.period.axis = list(at = seq(15)),
         main = paste("Wavelet coherence Germany, ", pm, " x detrended daily COVID-19 cases"))

# wc.sel.phases(wc, sel.period = 4, siglvl = 1, show.date = TRUE)
```



```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
## Wavelet coherence analysis on cluster averages
de_clstr = germany$de_clstr
for(c in unique(de_clstr$clstr_avg_detr$cluster_pm)){
  tmp = de_clstr$clstr_avg_detr[de_clstr$clstr_avg_detr$cluster_pm == c,]
  wc = compileWC(tmp)
  
  wc.image(wc, which.image = "wc",
           n.levels = 250, color.key = "interval",
           siglvl.contour = 0.1, siglvl.arrow = 0.1,
           legend.params = list(lab = "cross-wavelet power levels"), show.date = TRUE,
           spec.period.axis = list(at = seq(15)),
           main = paste("Germany, ", pm, " x detrended daily COVID-19 cases, cluster ", c))
  
  # wc.sel.phases(wc, sel.period = 4, siglvl = 1, show.date = TRUE)
}
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
# de_model_lag_gam = compileLaggedGAM(de_clstr)
# 
# ggplot(data = de_model_lag_gam[de_model_lag_gam$t < 5 & de_model_lag_gam$t > -5,], aes(x = as.factor(lag), y = t)) + 
#   geom_boxplot() + 
#   theme_bw() +
#   geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "red") + 
#   labs(title = paste("Dependence of individual COVID-19 time series and lagged ", pm, " concentrations"),x = "Time lag PM [days]", y = "t value of the partial coefficient")
# 
# de_model_geo = compileLaggedGLMGeo(de_model_lag_gam, germany$de_nuts3_map)

```