---
title: "Correlation of PM concentrations and SARS-CoV-2 infection"
always_allow_html: true
editor_options:
  chunk_output_type: console
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Set up working environment and defaults --------------------------------------
library(envimaR)
if(Sys.info()[["nodename"]] == "PC19616"){
  source("~/plygrnd/CovidAirPolution/CovidAirPolution/src/functions/000_setup.R")
} else {
  source("~/project/cov/CovidAirPolution/src/functions/000_setup.R")  
}
compile_data = FALSE

start_date = as.POSIXct("2020-01-15")
end_date = as.POSIXct("2020-04-20")
pm = "PM10"

Sys.setlocale("LC_TIME", "English") 

# Load Germany -----------------------------------------------------------------
if(pm == "PM2.5"){
  savefile = "germany_025.RDS"
} else {
  savefile = "germany_100.RDS"
}
if(compile_data){
  germany = compileDataDE(start_date = start_date, end_date = end_date,
                          pm = pm)
  saveRDS(germany, file.path(envrmt$path_tmp, savefile))
} else {
  germany = readRDS(file.path(envrmt$path_tmp, savefile))
}

# Compute additional datasets --------------------------------------------------
de_clstr = germany$de_clstr$clstr
de_model_lag = compileLaggedGLM(de_clstr)
de_model_geo = compileLaggedGLMGeo(de_model_lag, germany$de_nuts3_map)


ggplot(data = de_clstr, aes(x = date, y = cases, group = nuts3Code)) + geom_line()


start_date = lapply(unique(de_clstr$nuts3Code), function(n){
  tmp = de_clstr[de_clstr$nuts3Code == n, ]
  which(tmp$new_cases > 0)[1]
})
summary(unlist(start_date))

```




# Figure 1

## Figure 1a




## Figure 1a

c, Raw data on PM2.5 has been provided by the German environmental protection agency (UBN) and WAQI. ...

Data on total daily COVID-19 cases and deaths has been provided by the Robert-Koch-Institut. For some analysis, the data has been smoothed using a gaussian loess function with span 0.3 to reduce variation due to delayed reporting over the weekend and especcially on Sunday and Monday. The daily number of new infections has been computed from the daily variations of the total confirmed numbers.

The following figure shows the countrywide average of daily PM2.5 and smoothed COVID-19 new infections. Black vertical dotted lines represent the start of the contact restrictions (March 14th) and shut down (March 17th).
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
de_avg = germany$de_avg

tmp = de_avg[de_avg$date >= as.POSIXct("2020-03-01") & de_avg$date <= as.POSIXct("2020-04-01"),]
set.seed(01042020)
de_avg_glm_time = glm(new_cases ~ date + weekday , family = quasipoisson, data = tmp)
tmp$predicted = predict(de_avg_glm_time, tmp, type = "response")
tmp$residuals = residuals(de_avg_glm_time, type="response")

figure_1a = ggplot() + 
  geom_line(data = de_avg, aes(x = date, y = pm_mean, color = "Daily mean PM10")) +
  geom_point(data = de_avg, aes(x = date, y = pm_mean, color = "Daily mean PM10")) + 
  geom_line(data = de_avg, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_point(data = de_avg, aes(x = date, y = new_cases, color = "Daily new cases")) + 
  geom_line(data = tmp, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
  geom_point(data = tmp, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
  geom_line(data = tmp, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_point(data = tmp, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype="dotted", color = "black") +
  geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype="dotted", color = "black") +
  # geom_text(aes(x = as.POSIXct("2020-03-14"), label="Start of contact restrictions \n", y=-1), colour="black", angle=90, size = 3) +
  # geom_text(aes(x = as.POSIXct("2020-03-17"), label="Start of shut down \n", y=-1), colour="black", angle=90, size = 3) +
  labs(x = "Date and day of week", y = expression("Infections, PM10 ~µm/m^{3}")) +
  theme_bw() + 
  scale_x_datetime(date_labels = "%d.%m, %a", date_breaks = "2 day", date_minor_breaks = "1 day") + 
  scale_color_manual(values=c("#b2df8a", "#1f78b4", "#a6cee3")) +
  theme(text = element_text(size = 10), axis.title = element_text(size = 10),
        legend.position = c(0.35, 0.75), axis.text.x = element_text(angle=90, hjust=0.95, vjust=0.2), 
        legend.title = element_blank(), panel.background = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# ggsave(file.path(envrmt$path_figures, "figure_1b.png"), plot = figure_1b,
#        dpi = 300)
# 
# print(figure_1b)
```


## Figure 1b

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
lims <- as.POSIXct(c(min(de_avg$date), max(de_avg$date)))

figure_1b = ggplot() + 
  geom_line(data = tmp, aes(x = date, y = residuals, color = "Daily new cases, residuals")) +
  geom_point(data = tmp, aes(x = date, y = residuals, color = "Daily new cases, residuals")) +
  geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype="dotted", color = "black") +
  geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype="dotted", color = "black") +
  geom_text(aes(x = as.POSIXct("2020-03-14"), label="Start of contact restrictions \n", y=-4), colour="black", angle=90, size = 2) +
  geom_text(aes(x = as.POSIXct("2020-03-17"), label="Start of shut down \n", y=-4), colour="black", angle=90, size = 2) +
  labs(x = "Date and day of week", y = "Residuals") +
  theme_bw() + 
  scale_x_datetime(date_labels = "%d.%m, %a", date_breaks = "2 day", date_minor_breaks = "1 day", limits = lims) +
  scale_color_manual(values=c("#d95f02")) +
  theme(text = element_text(size = 10), axis.title = element_text(size = 10),
        legend.position = "none", axis.text.x = element_text(angle=90, hjust=0.95, vjust=0.2), 
        legend.title = element_blank(), panel.background = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# ggsave(file.path(envrmt$path_figures, "figure_1c.png"), plot = figure_1c,
#        dpi = 300)
# 
# print(figure_1c)
```



## Figure 1 combined

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
# figure_1 = plot_grid(figure_1a, figure_1b, labels = c("b", "c"), ncol = 1)
figure_1 = plot_grid(figure_1a, figure_1b, labels = "auto", label_size = 10, ncol = 1)

ggsave(file.path(envrmt$path_figures, "figure_1.png"), plot = figure_1, 
       width = 16, height = 16, units = "cm", dpi = 300)
```


# Figure 2

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
de_model_lag = compileLaggedGLM(data = de_clstr,  
                                pm = "org", 
                                frml = "new_cases ~ date + weekday + pm_mean_lag",
                                nlags = 14,
                                subset_var = "new_cases", subset_thv = 1, individual = "start", ndays = c(0, 30))


t_matrix = reshape(de_model_lag[, c("nuts3Code", "t", "lag")], idvar = "nuts3Code", timevar = "lag", direction = "wide")
rownames(t_matrix) = t_matrix$nuts3Code
vars = c("pm_mean", "pop_total", "pop_dens", "centroid_lat", "centroid_lon", "st_area")
t_environment = aggregate(st_drop_geometry(de_clstr[, vars]), by = list(de_clstr$nuts3Code), FUN = mean)
names(t_environment)[1] = "nuts3Code"
t_environment = merge(t_environment, info)
t_matrix = merge(t_matrix, t_environment, by = "nuts3Code")
vars = c(vars, "date_start", "first_shutdown", "max_shutdown", "cases_at_max", "daily_max")
rd = rda(t_matrix[, 2:22], t_matrix[, c("pm_mean", "pop_dens", "first_shutdown",
                                        "cases_at_max", "centroid_lat")])

plot(rd, type = "points", scaling = 0)


figure_2a = ggplot(data = de_model_lag, aes(x = as.factor(lag), y = t)) + 
  geom_boxplot() + 
  theme_bw() +
  geom_hline(yintercept = c(0), linetype="dashed", color = "black") + 
  scale_y_continuous(limits = c(-8, 12), breaks = seq(-10, 20, 2)) +
  labs(x = "Time lag [days]", y = "Effect size") + 
  theme(text = element_text(size = 10), axis.title = element_text(size = 10),
        panel.background = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

ggsave(file.path(envrmt$path_figures, "figure_2a.png"), plot = figure_2a, 
       width = 16, height = 8, units = "cm", dpi = 300)

figure_2b = ggplot(data = de_model_geo$test_geo_lag) + 
  geom_line(aes(x = lag, y = centroid_lat_t, color = "Latitude")) +
  geom_point(aes(x = lag, y = centroid_lat_t, color = "Latitude"), size = 2) +
  geom_line(aes(x = lag, y = centroid_lon_t, color = "Longitude")) +
  geom_point(aes(x = lag, y = centroid_lon_t, color = "Longitude"), size = 2) +
  geom_line(aes(x = lag, y = area_t, color = "Area")) +
  geom_point(aes(x = lag, y = area_t, color = "Area"), size = 2) +
  scale_x_continuous(breaks = seq(-15, 0, 1), minor_breaks = seq(-15, 0, 1)) +
  theme_bw() + 
  scale_color_manual(values=c('grey', 'red',"darkblue")) +
  scale_y_continuous(limits = c(-8, 12), breaks = seq(-10, 20, 2)) +
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "black") + 
  labs(x = "Time lag [days]", y = " ", 
       color = "") + 
  theme(legend.position = c(0.8, 0.8), panel.background = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())


figure_2c = ggplot(data = de_model_geo$test_geo[de_model_geo$test_geo$lag == -12, ], 
                   aes(x = centroid_lat, y = t)) + 
  geom_point(color = "red", size = 2) +
  theme_bw() + 
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "black") + 
  scale_x_continuous(breaks = seq(47, 55, 1)) + 
  scale_y_continuous(limits = c(-7, 4), breaks = seq(-10, 20, 2)) +
  labs(x = "Latitude [°N]", y = "Effect size", 
       fill = "", shape = "", size = "") + 
  theme(legend.position = c(0.65, 0.08), legend.direction="horizontal", 
        panel.background = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())


figure_2d = ggplot(data = de_model_geo$test_geo[de_model_geo$test_geo$lag == -6, ], 
                   aes(x = centroid_lat, y = t)) + 
  geom_point(color = "red", size = 2) +
  theme_bw() + 
  geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "black") + 
  scale_x_continuous(breaks = seq(47, 55, 1)) + 
  scale_y_continuous(limits = c(-7, 4), breaks = seq(-10, 20, 2)) +
  labs(x = "Latitude [°N]", y = " ", 
       fill = "", shape = "", size = "") + 
  theme(legend.position = c(0.65, 0.08), legend.direction="horizontal", 
        panel.background = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())


figure_2 = plot_grid(figure_2a, figure_2b, figure_2c, figure_2d, labels = "auto")
ggsave(file.path(envrmt$path_figures, "figure_2.png"), plot = figure_2, dpi = 300)
```

Dependence of individual COVID-19 time series and lagged PM10 concentrations.
effect size = t-value of the partial coefficient.




# Supplement

## Maps
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

background_map = st_transform(ne_countries(scale = "large", continent = "europe", returnclass = "sf"),
                              crs = "+init=epsg:25832")
layer = st_transform(germany$de_nuts3_map, crs = "+init=epsg:25832")

figure_2e = ggplot(data = background_map) +
  geom_sf() + 
  geom_sf(data = layer, aes(fill = cases)) +
  scale_fill_viridis_c(trans = "log10") +
  theme(legend.position="bottom") +
  labs(fill = "Infections (01.04.2020)") +
  coord_sf(xlim = c((st_bbox(layer)["xmin"] - 10000), 
                    (st_bbox(layer)["xmax"] + 10000)), 
           ylim = c((st_bbox(layer)["ymin"] - 25000),
                    (st_bbox(layer)["ymax"] + 10000)), expand = FALSE) + 
  annotation_north_arrow(location = "tl", which_north = "true",
                         height = unit(0.75, "cm"), width = unit(0.75, "cm"),
                         pad_x = unit(0.0, "in"), pad_y = unit(0.1, "in"),
                         style = north_arrow_minimal)

ggsave(file.path(envrmt$path_figures, "figure_2e.png"), plot = figure_2e, 
       width = 16, height = 16, units = "cm", dpi = 300)

background_map = st_transform(ne_countries(scale = "large", continent = "europe", returnclass = "sf"),
                              crs = "+init=epsg:25832")

l = -4
p = 0.1

layer1 = st_set_geometry(de_model_geo$test_geo[de_model_geo$test_geo$lag == l &
                                                de_model_geo$test_geo$p < p, ], 
                        de_model_geo$test_geo$geometry[de_model_geo$test_geo$lag == l &
                                                de_model_geo$test_geo$p < p, ])

layer1 = st_set_geometry(de_model_geo$test_geo[de_model_geo$test_geo$p < p, ], 
                        de_model_geo$test_geo$geometry[de_model_geo$test_geo$p < p, ])


layer1 = st_set_geometry(de_model_geo$test_geo[de_model_geo$test_geo$t > 0 &
                                                de_model_geo$test_geo$p < p, ], 
                        de_model_geo$test_geo$geometry[de_model_geo$test_geo$t > 0 &
                                                de_model_geo$test_geo$p < p, ])


layer1 = st_transform(layer1, crs = "+init=epsg:25832")

figure_2f = ggplot(data = background_map) +
  geom_sf() + 
  geom_sf(data = layer) +
  geom_sf(data = layer1, aes(fill = lag)) +
  scale_fill_viridis_c() +
  theme(legend.position="bottom") +
  labs(fill = "Maximum lag of significant positive effect size") +
  coord_sf(xlim = c((st_bbox(layer)["xmin"] - 10000), 
                    (st_bbox(layer)["xmax"] + 10000)), 
           ylim = c((st_bbox(layer)["ymin"] - 25000),
                    (st_bbox(layer)["ymax"] + 10000)), expand = FALSE) + 
  annotation_north_arrow(location = "tl", which_north = "true",
                         height = unit(0.5, "cm"), width = unit(0.5, "cm"),
                         pad_x = unit(0.0, "in"), pad_y = unit(0.1, "in"),
                         style = north_arrow_minimal)

ggsave(file.path(envrmt$path_figures, "figure_2f.png"), plot = figure_2f, 
       width = 16, height = 16, units = "cm", dpi = 300)

figure_2_2 = plot_grid(figure_2e, figure_2f, labels = c("e", "f"))

```


## Figure: Smooth and PM2.5

First row: Boxplot for PM10_new_cases_smooth, Latitude-Lag-Plot smooth
Second row: Boxplot for PM2.5, Latitude-Lag-Plot 2.5


<!-- figure_2c = ggplot() +  -->
<!--   geom_point(data = de_model_geo$test_geo[de_model_geo$test_geo$lag <= -10, ],  -->
<!--              aes(x = centroid_lat, y = t, fill = as.factor(lag),  -->
<!--                  shape = as.factor(lag), size = as.factor(lag))) + -->
<!--   scale_shape_manual(values = c(22, 23, 21, 24, 25)) + -->
<!--   scale_fill_manual(values = c('grey', 'grey','red', 'grey', 'grey')) + -->
<!--   scale_size_manual(values = c(1.5, 1.5, 3, 1.5, 1.5)) +  -->
<!--   geom_smooth(data = de_model_geo$test_geo[de_model_geo$test_geo$lag == -12, ],  -->
<!--               aes(x = centroid_lat, y = t, ), method = "lm", color = "red")+ -->
<!--   theme_bw() +  -->
<!--   geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "black") +  -->
<!--   scale_x_continuous(breaks = seq(47, 55, 1)) +  -->
<!--   scale_y_continuous(breaks = seq(-10, 20, 2)) + -->
<!--   labs(x = "Latitude (degrees)", y = "Effect size",  -->
<!--        fill = "", shape = "", size = "") +  -->
<!--   theme(legend.position = c(0.65, 0.08), legend.direction="horizontal",  -->
<!--         panel.background = element_blank(), panel.grid.major = element_blank(), -->
<!--         panel.grid.minor = element_blank()) -->


<!-- figure_2d = ggplot() +  -->
<!--   geom_point(data = de_model_geo$test_geo[de_model_geo$test_geo$lag >= -8 & -->
<!--                                             de_model_geo$test_geo$lag <= -3, ],  -->
<!--              aes(x = centroid_lat, y = t, fill = as.factor(lag),  -->
<!--                  shape = as.factor(lag), size = as.factor(lag))) + -->
<!--   scale_shape_manual(values = c(22, 23, 21, 24, 25, 21)) + -->
<!--   scale_fill_manual(values = c('grey', 'grey','red', 'grey', 'grey', "blue")) + -->
<!--   scale_size_manual(values = c(1.5, 1.5, 3, 1.5, 1.5, 3)) +  -->
<!--   geom_smooth(data = de_model_geo$test_geo[de_model_geo$test_geo$lag == -6, ],  -->
<!--               aes(x = centroid_lat, y = t, ), method = "lm", color = "red")+ -->
<!--   theme_bw() +  -->
<!--   geom_hline(yintercept = c(-2, 2), linetype="dashed", color = "black") +  -->
<!--   scale_x_continuous(breaks = seq(47, 55, 1)) +  -->
<!--   scale_y_continuous(breaks = seq(-10, 20, 2)) + -->
<!--   labs(x = "Latitude (degrees)", y = " ",  -->
<!--        fill = "", shape = "", size = "") +  -->
<!--   theme(legend.position = c(0.3, 0.09), legend.direction="horizontal",  -->
<!--         panel.background = element_blank(), panel.grid.major = element_blank(), -->
<!--         panel.grid.minor = element_blank()) -->

