---
title: "Media for the publication on the correlation between PM and SARS-CoV-2 infections in Germany"
author: "Thomas Nauss et al."
always_allow_html: true
editor_options:
  chunk_output_type: console
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# Set up working environment and load graphics.
library(envimaR)
if (Sys.info()[["nodename"]] == "PC19616") {
  source("~/plygrnd/CovidAirPolution/CovidAirPolution/src/functions/000_setup.R")
} else {
  source("~/project/cov/CovidAirPolution/src/functions/000_setup.R")
}

pm_vars <- c("PM2.5", "PM10")
de_cmpldata <- lapply(pm_vars, function(pm) {
  cmpldata_file <- paste0("germany_", pm, "_extended.RDS")
  return(readRDS(file.path(envrmt$path_analysis, cmpldata_file)))
})
names(de_cmpldata) <- pm_vars
de_figure_cntry_avg <- readRDS(file.path(envrmt$path_figures, "germany_figure_cntry_avg.rds"))
de_figure_gam_lag_set <- readRDS(file.path(envrmt$path_figures, "germany_figure_gam_lag_set.rds"))
de_figure_gam_lag_mixed_set <- readRDS(file.path(envrmt$path_figures, "germany_figure_gam_lag_mixed_set.rds"))
de_model_figure_cumulative_effect <- readRDS(file.path(envrmt$path_figures, "germany_model_figure_cumulative_effect.rds"))
de_map_covid_infections <- readRDS(file.path(envrmt$path_figures, "germany_map_covid_infections.rds"))
de_map_pm_mean <- readRDS(file.path(envrmt$path_figures, "germany_map_pm_mean.rds"))
de_cor_figure_corr_PM10_PM2.5 <- readRDS(file.path(envrmt$path_figures, "germany_cor_figure_corr_PM10_PM2.5.rds"))
de_excluded_data_PM10 <- readRDS(file.path(envrmt$path_analysis, "germany_PM10_non_valid.rds"))
de_excluded_data_PM2.5 <- readRDS(file.path(envrmt$path_analysis, "germany_PM2.5_non_valid.rds"))

lag_var <- "pm_mean_estm_best"
```



# Statistical information included in the manuscript.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

mean(de_model_figure_cumulative_effect$PM10$add_info$difftime_start_max)
quantile(de_model_figure_cumulative_effect$PM10$add_info$difftime_start_max)
sum(de_model_figure_cumulative_effect$PM10$add_info$daily_max >= 10)

mean(de_model_figure_cumulative_effect$PM2.5$add_info$difftime_start_max)
quantile(de_model_figure_cumulative_effect$PM2.5$add_info$difftime_start_max)
sum(de_model_figure_cumulative_effect$PM2.5$add_info$daily_max >= 10)

de_excluded_data_PM10
de_cor_figure_corr_PM10_PM2.5$n_PM10_nuts3_regions
de_cor_figure_corr_PM10_PM2.5$figure_PM10_outliers_nbrs

de_excluded_data_PM2.5
de_cor_figure_corr_PM10_PM2.5$n_PM2.5_nuts3_regions
de_cor_figure_corr_PM10_PM2.5$figure_PM2.5_outliers_nbrs

de_cor_figure_corr_PM10_PM2.5$n_PM10_PM2.5_nuts3_regions_overlap
de_cor_figure_corr_PM10_PM2.5$cntry_indv_ovrlp_cor


# Districts exceeding WHO PM10 threshold value
cntry_indv_10 <- de_cmpldata$PM10$de_clstr$clstr
length(sort(unique(cntry_indv_10$nuts3Code[cntry_indv_10$pm_mean_estm_best > 50])))

# Number of days with exceeding WHO PM10 threshold value by district
table(sort(cntry_indv_10$nuts3Code[cntry_indv_10$pm_mean_estm_best > 50]))
min(table(sort(cntry_indv_10$nuts3Code[cntry_indv_10$pm_mean_estm_best > 50])))
max(table(sort(cntry_indv_10$nuts3Code[cntry_indv_10$pm_mean_estm_best > 50])))

# Total study days exceeding WHO PM10 threshold value
sum(cntry_indv_10$pm_mean_estm_best > 50)
signif(sum(cntry_indv_10$pm_mean_estm_best > 50) / nrow(cntry_indv_10) * 100, 2)

# Unique days exceeding WHO PM10 threshold value
sort(unique(cntry_indv_10$date[cntry_indv_10$pm_mean_estm_best > 50]))


# Districts exceeding WHO PM2.5 threshold value
cntry_indv_25 <- de_cmpldata$PM2.5$de_clstr$clstr
length(sort(unique(cntry_indv_25$nuts3Code[cntry_indv_25$pm_mean_estm_best > 25])))

# Number of days with exceeding WHO PM2.5 threshold value by district
table(sort(cntry_indv_25$nuts3Code[cntry_indv_25$pm_mean_estm_best > 25]))
min(table(sort(cntry_indv_25$nuts3Code[cntry_indv_25$pm_mean_estm_best > 25])))
max(table(sort(cntry_indv_25$nuts3Code[cntry_indv_25$pm_mean_estm_best > 25])))

# Total study days exceeding WHO PM2.5 threshold value
sum(cntry_indv_25$pm_mean_estm_best > 25)
signif(sum(cntry_indv_25$pm_mean_estm_best > 25) / nrow(cntry_indv_25) * 100, 2)

# Unique days exceeding WHO PM10 threshold value
sort(unique(cntry_indv_25$date[cntry_indv_25$pm_mean_estm_best > 25]))
```



# Figure Part 1: Country wide reported SARS-CoV-2 infections and PM10 concentrations.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_1_1 <- plot_grid(de_figure_cntry_avg$PM10$cntry_avg,
  de_figure_cntry_avg$PM10$cntry_avg_pm,
  de_figure_cntry_avg$PM10$cntry_avg_covid,
  labels = c("A", "B", "C"), label_size = 12, ncol = 1
)

figure_1_2 <- plot_grid(de_map_covid_infections$PM10, de_map_pm_mean$PM10, labels = c("d", "e"), label_size = 12, ncol = 2)

figure_1 <- plot_grid(figure_1_1, figure_1_2, labels = NULL, ncol = 1)
# ggsave(file.path(envrmt$path_figures, "figure_1_germany_PM10.png"),
#   plot = figure_1,
#   width = 347, height = 450, units = "mm", dpi = 300
# )
```



# Figure Part 2: Country wide SARS-CoV-2 infection models for PM10.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_2_1 <- plot_grid(de_model_figure_cumulative_effect$PM10$cumulative_effect_log,
  de_model_figure_cumulative_effect$PM10$cumulative_effect,
  labels = c("F", "G"), label_size = 10, ncol = 1
)

figure_2_2 <- plot_grid(de_figure_gam_lag_set$PM10$pm_mean_estm_best, de_figure_gam_lag_mixed_set$PM10$pm_mean_estm_best$estimate_mixed,
  labels = c("H", "I"), label_size = 10, ncol = 2
)

figure_2 <- plot_grid(figure_2_1, figure_2_2, labels = NULL, ncol = 1)
# ggsave(file.path(envrmt$path_figures, "figure_2_germany_PM10.png"),
#   plot = figure_2,
#   width = 347, height = 450, units = "mm", dpi = 300
# )
```




# Figure: Combination of Figure Part 1 and 2.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure <- plot_grid(figure_1, figure_2, labels = NULL, ncol = 2)
ggsave(file.path(envrmt$path_figures, "figure_germany_PM10.png"),
  plot = figure,
  width = 694, height = 450, units = "mm", dpi = 300
)
```




# Supplementary Figure: Country wide reported SARS-CoV-2 infections and PM2.5 concentrations.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_sup_1_1 <- plot_grid(de_figure_cntry_avg$PM2.5$cntry_avg,
  de_figure_cntry_avg$PM2.5$cntry_avg_pm,
  de_figure_cntry_avg$PM2.5$cntry_avg_covid,
  labels = c("A", "B", "C"), label_size = 12, ncol = 1
)

figure_sup_1_2 <- plot_grid(de_map_covid_infections$PM2.5, de_map_pm_mean$PM2.5, labels = c("D", "E"), label_size = 12, ncol = 2)

figure_sup_1 <- plot_grid(figure_sup_1_1, figure_sup_1_2, labels = NULL, ncol = 1)
ggsave(file.path(envrmt$path_figure, "figure_s1_germany_PM25.png"),
  plot = figure_sup_1,
  width = 347, height = 450, units = "mm", dpi = 300
)
```




# Supplementary Figure: Country wide SARS-CoV-2 infection models for PM2.5.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_sup_2_1 <- plot_grid(de_model_figure_cumulative_effect$PM2.5$cumulative_effect_log,
  de_model_figure_cumulative_effect$PM2.5$cumulative_effect,
  labels = c("A", "B"), label_size = 10, ncol = 1
)

figure_sup_2_2 <- plot_grid(de_figure_gam_lag_set$PM2.5$pm_mean_estm_best, de_figure_gam_lag_mixed_set$PM2.5$pm_mean_estm_best$estimate_mixed,
  labels = c("C", "D"), label_size = 10, ncol = 2
)

figure_sup_2 <- plot_grid(figure_sup_2_1, figure_sup_2_2, labels = NULL, ncol = 1)
ggsave(file.path(envrmt$path_figures, "figure_s2_germany_PM25.png"),
  plot = figure_sup_2,
  width = 347, height = 450, units = "mm", dpi = 300
)
```



# Supplementary Figure: Correlation between PM10 and PM2.5.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_sup_3_1 <- plot_grid(de_cor_figure_corr_PM10_PM2.5$figure_corr_PM10_PM2.5,
  de_cor_figure_corr_PM10_PM2.5$figure_PM10_outliers,
  de_cor_figure_corr_PM10_PM2.5$figure_PM2.5_outliers,
  labels = c("A", "B", "C"), label_size = 12, ncol = 3
)

figure_sup_3_2 <- plot_grid(de_figure_gam_lag_set$PM10$pm_mean, de_figure_gam_lag_mixed_set$PM10$pm_mean$estimate_mixed,
  labels = c("D", "E"), label_size = 10, ncol = 2
)

figure_sup_3_3 <- plot_grid(de_figure_gam_lag_set$PM2.5$pm_mean, de_figure_gam_lag_mixed_set$PM2.5$pm_mean$estimate_mixed,
  labels = c("F", "G"), label_size = 10, ncol = 2
)

figure_sup_3 <- plot_grid(figure_sup_3_1, figure_sup_3_2, figure_sup_3_3, labels = NULL, ncol = 1)

ggsave(file.path(envrmt$path_figures, "figure_s3_germany_mixed.png"),
  plot = figure_sup_3,
  width = 347, height = 450, units = "mm", dpi = 300
)
```




# Table S1: Analysis of the number of infections on April 1st 2020 with regard to PM10 concentrations.
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

de_model_figure_cumulative_effect$PM10$lm_tavrg_smry
signif(de_model_figure_cumulative_effect$PM10$lm_tavrg_smry$coefficients, 3)
signif(de_model_figure_cumulative_effect$PM10$lm_tavrg_smry$adj.r.squared, 2)
```




# Table S2: Final model from the analysis of the number of infections on April 1st 2020 with regard to PM10 concentrations.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

de_model_figure_cumulative_effect$PM10$lm_tavrg_stpAIC_smry
signif(de_model_figure_cumulative_effect$PM10$lm_tavrg_stpAIC_smry$coefficients, 3)
signif(de_model_figure_cumulative_effect$PM10$lm_tavrg_stpAIC_smry$adj.r.squared, 2)
```



# Table S3: Analysis of the number of infections on April 1st 2020 with regard to PM2.5 concentrations.
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

de_model_figure_cumulative_effect$PM2.5$lm_tavrg_smry
signif(de_model_figure_cumulative_effect$PM2.5$lm_tavrg_smry$coefficients, 3)
signif(de_model_figure_cumulative_effect$PM2.5$lm_tavrg_smry$adj.r.squared, 2)
```



# Table S4: Final model from the analysis of the number of infections on April 1st 2020 with regard to PM2.5 concentrations.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

de_model_figure_cumulative_effect$PM2.5$lm_tavrg_stpAIC_smry
signif(de_model_figure_cumulative_effect$PM2.5$lm_tavrg_stpAIC_smry$coefficients, 3)
signif(de_model_figure_cumulative_effect$PM2.5$lm_tavrg_stpAIC_smry$adj.r.squared, 2)
```
