---
title: "Media for the publication on the correlation between PM and SARS-CoV-2 infections in Germany"
author: "Thomas Nauss et al."
always_allow_html: true
editor_options:
  chunk_output_type: console
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# Set up working environment and load graphics.
library(envimaR)
if (Sys.info()[["nodename"]] == "PC19616") {
  source("~/plygrnd/CovidAirPolution/CovidAirPolution/src/functions/000_setup.R")
} else {
  source("~/project/cov/CovidAirPolution/src/functions/000_setup.R")
}

pm_vars <- c("PM2.5", "PM10")
cmpldata <- lapply(pm_vars, function(pm) {
  cmpldata_file <- paste0("italy_", pm, "_extended.RDS")
  return(readRDS(file.path(envrmt$path_analysis, cmpldata_file)))
})
names(cmpldata) <- pm_vars
figure_cntry_avg <- readRDS(file.path(envrmt$path_figures, "italy_figure_cntry_avg.rds"))
figure_gam_lag_set <- readRDS(file.path(envrmt$path_figures, "italy_figure_gam_lag_set.rds"))
figure_gam_lag_mixed_set <- readRDS(file.path(envrmt$path_figures, "italy_figure_gam_lag_mixed_set.rds"))
model_figure_cumulative_effect <- readRDS(file.path(envrmt$path_figures, "italy_model_figure_cumulative_effect.rds"))
map_covid_infections <- readRDS(file.path(envrmt$path_figures, "italy_map_covid_infections.rds"))
map_pm_mean <- readRDS(file.path(envrmt$path_figures, "italy_map_pm_mean.rds"))
# lm_tavrg_smry <- readRDS(file.path(envrmt$path_figures, "italy_lm_tavrg_smry.rds"))
# lm_tavrg_stpAIC_smry <- readRDS(file.path(envrmt$path_figures, "italy_lm_tavrg_stpAIC_smry.rds"))
cor_figure_corr_PM10_PM2.5 <- readRDS(file.path(envrmt$path_figures, "italy_cor_figure_corr_PM10_PM2.5.rds"))
excluded_data_PM10 <- readRDS(file.path(envrmt$path_analysis, "italy_PM10_non_valid.rds"))
excluded_data_PM2.5 <- readRDS(file.path(envrmt$path_analysis, "italy_PM2.5_non_valid.rds"))

lag_var <- "pm_mean_estm_best"
```



# Statistical information included in the manuscript.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

mean(model_figure_cumulative_effect$PM10$add_info$difftime_start_max)
quantile(model_figure_cumulative_effect$PM10$add_info$difftime_start_max)
sum(model_figure_cumulative_effect$PM10$add_info$daily_max >= 10)

mean(model_figure_cumulative_effect$PM2.5$add_info$difftime_start_max)
quantile(model_figure_cumulative_effect$PM2.5$add_info$difftime_start_max)
sum(model_figure_cumulative_effect$PM2.5$add_info$daily_max >= 10)

excluded_data_PM10
cor_figure_corr_PM10_PM2.5$n_PM10_nuts3_regions
cor_figure_corr_PM10_PM2.5$figure_PM10_outliers_nbrs

excluded_data_PM2.5
cor_figure_corr_PM10_PM2.5$n_PM2.5_nuts3_regions
cor_figure_corr_PM10_PM2.5$figure_PM2.5_outliers_nbrs

cor_figure_corr_PM10_PM2.5$n_PM10_PM2.5_nuts3_regions_overlap
cor_figure_corr_PM10_PM2.5$cntry_indv_ovrlp_cor


# Districts exceeding WHO PM10 threshold value
cntry_indv <- cmpldata$PM10$it_clstr$clstr
length(sort(unique(cntry_indv$nuts3Code[cntry_indv$pm_mean_estm_best > 50])))

# Number of days with exceeding WHO PM10 threshold value by district
table(sort(cntry_indv$nuts3Code[cntry_indv$pm_mean_estm_best > 50]))
max(table(sort(cntry_indv$nuts3Code[cntry_indv$pm_mean_estm_best > 50])))

# Total study days exceeding WHO PM10 threshold value
sum(cntry_indv$pm_mean_estm_best > 50)
sum(cntry_indv$pm_mean_estm_best > 50) / nrow(cntry_indv) * 100

# Unique days exceeding WHO PM10 threshold value
sort(unique(cntry_indv$date[cntry_indv$pm_mean_estm_best > 50]))

# Districts with at least 10 new infections per day once during the study (PM10 sample)
length(unique(cntry_indv$nuts3Code[cntry_indv$new_cases >= 10]))


# Districts exceeding WHO PM2.5 threshold value
cntry_indv <- cmpldata$PM2.5$it_clstr$clstr
length(sort(unique(cntry_indv$nuts3Code[cntry_indv$pm_mean_estm_best > 25])))

# Number of days with exceeding WHO PM2.5 threshold value by district
table(sort(cntry_indv$nuts3Code[cntry_indv$pm_mean_estm_best > 25]))
max(table(sort(cntry_indv$nuts3Code[cntry_indv$pm_mean_estm_best > 25])))

# Total study days exceeding WHO PM2.5 threshold value
sum(cntry_indv$pm_mean_estm_best > 25)
sum(cntry_indv$pm_mean_estm_best > 25) / nrow(cntry_indv) * 100

# Unique days exceeding WHO PM2.5 threshold value
sort(unique(cntry_indv$date[cntry_indv$pm_mean_estm_best > 25]))

# Districts with at least 10 new infections per day once during the study (PM2.5 sample)
length(unique(cntry_indv$nuts3Code[cntry_indv$new_cases >= 10]))

for(i in unique(cntry_indv$nuts3Code)){
  print(cntry_indv$date[cntry_indv$nuts3Code == i])
}

```



# Fig. 1: Country wide reported SARS-CoV-2 infections and PM10 concentrations.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_1_1 <- plot_grid(figure_cntry_avg$PM10$cntry_avg,
  figure_cntry_avg$PM10$cntry_avg_pm,
  figure_cntry_avg$PM10$cntry_avg_covid,
  labels = c("A", "B", "C"), label_size = 12, ncol = 1
)

figure_1_2 <- plot_grid(map_covid_infections$PM10, map_pm_mean$PM10, labels = c("D", "E"), label_size = 12, ncol = 2)

figure_1 <- plot_grid(figure_1_1, figure_1_2, labels = NULL, ncol = 1)
ggsave(file.path(envrmt$path_figures, "italy_figure_s4.png"),
  plot = figure_1,
  width = 347, height = 450, units = "mm", dpi = 300
)
```



# Fig. 2: Country wide reported SARS-CoV-2 infections and PM10 concentrations.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_2_1 <- plot_grid(model_figure_cumulative_effect$PM10$cumulative_effect_log,
  model_figure_cumulative_effect$PM10$cumulative_effect,
  labels = c("A", "B"), label_size = 10, ncol = 1
)

figure_2_2 <- plot_grid(figure_gam_lag_set$PM10$pm_mean_estm_best, figure_gam_lag_mixed_set$PM10$pm_mean_estm_best$estimate_mixed,
  labels = c("C", "D"), label_size = 10, ncol = 2
)

figure_2 <- plot_grid(figure_2_1, figure_2_2, labels = NULL, ncol = 1)
ggsave(file.path(envrmt$path_figures, "italy_figure_s5.png"),
  plot = figure_2,
  width = 347, height = 450, units = "mm", dpi = 300
)
```



# Extended Data Table 1: Analysis of the number of infections on April 1st 2020 with regard to PM10 concentrations.
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

model_figure_cumulative_effect$PM10$lm_tavrg_smry
signif(model_figure_cumulative_effect$PM10$lm_tavrg_smry$coefficients, 3)
```



# Supplementary Fig. 1: Country wide reported SARS-CoV-2 infections and PM2.5 concentrations.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_sup_1_1 <- plot_grid(figure_cntry_avg$PM2.5$cntry_avg,
  figure_cntry_avg$PM2.5$cntry_avg_pm,
  figure_cntry_avg$PM2.5$cntry_avg_covid,
  labels = c("A", "B", "C"), label_size = 12, ncol = 1
)

figure_sup_1_2 <- plot_grid(map_covid_infections$PM2.5, map_pm_mean$PM2.5, labels = c("D", "E"), label_size = 12, ncol = 2)

figure_sup_1 <- plot_grid(figure_sup_1_1, figure_sup_1_2, labels = NULL, ncol = 1)
ggsave(file.path(envrmt$path_figure, "italy_figure_s6.png"),
  plot = figure_sup_1,
  width = 347, height = 450, units = "mm", dpi = 300
)
```




# Supplementary Fig. 2: Country wide reported SARS-CoV-2 infections and PM10 concentrations.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_sup_2_1 <- plot_grid(model_figure_cumulative_effect$PM2.5$cumulative_effect_log,
  model_figure_cumulative_effect$PM2.5$cumulative_effect,
  labels = c("A", "B"), label_size = 10, ncol = 1
)

figure_sup_2_2 <- plot_grid(figure_gam_lag_set$PM2.5$pm_mean_estm_best, figure_gam_lag_mixed_set$PM2.5$pm_mean_estm_best$estimate_mixed,
  labels = c("C", "D"), label_size = 10, ncol = 2
)

figure_sup_2 <- plot_grid(figure_sup_2_1, figure_sup_2_2, labels = NULL, ncol = 1)
ggsave(file.path(envrmt$path_figures, "italy_figure_s7.png"),
  plot = figure_sup_2,
  width = 347, height = 450, units = "mm", dpi = 300
)
```



# Supplementary Fig. 3: Correlation between PM10 and PM2.5.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_sup_3_1 <- plot_grid(cor_figure_corr_PM10_PM2.5$figure_corr_PM10_PM2.5,
  cor_figure_corr_PM10_PM2.5$figure_PM10_outliers,
  cor_figure_corr_PM10_PM2.5$figure_PM2.5_outliers,
  labels = c("A", "B", "C"), label_size = 12, ncol = 3
)

figure_sup_3_2 <- plot_grid(figure_gam_lag_set$PM10$pm_mean, figure_gam_lag_mixed_set$PM10$pm_mean$estimate_mixed,
  labels = c("D", "E"), label_size = 10, ncol = 2
)

figure_sup_3_3 <- plot_grid(figure_gam_lag_set$PM2.5$pm_mean, figure_gam_lag_mixed_set$PM2.5$pm_mean$estimate_mixed,
  labels = c("F", "G"), label_size = 10, ncol = 2
)

figure_sup_3 <- plot_grid(figure_sup_3_1, figure_sup_3_2, figure_sup_3_3, labels = NULL, ncol = 1)

ggsave(file.path(envrmt$path_figures, "italy_figure_s8.png"),
  plot = figure_sup_3,
  width = 347, height = 450, units = "mm", dpi = 300
)
```



# Supplementary Table 1: Final model from the analysis of the number of infections on April 1st 2020 with regard to PM10 concentrations.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

model_figure_cumulative_effect$PM10$lm_tavrg_stpAIC_smry
signif(model_figure_cumulative_effect$PM10$lm_tavrg_stpAIC_smry$coefficients, 3)
```



# Supplementary Table 2: Analysis of the number of infections on April 1st 2020 with regard to PM2.5 concentrations.
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

model_figure_cumulative_effect$PM2.5$lm_tavrg_smry
signif(model_figure_cumulative_effect$PM2.5$lm_tavrg_smry$coefficients, 3)
```



# Supplementary Table 3: Final model from the analysis of the number of infections on April 1st 2020 with regard to PM2.5 concentrations.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

model_figure_cumulative_effect$PM2.5$lm_tavrg_stpAIC_smry
signif(model_figure_cumulative_effect$PM2.5$lm_tavrg_stpAIC_smry$coefficients, 3)

```
