---
title: "Media for the publication on the correlation between PM and SARS-CoV-2 infections in Germany"
author: "Thomas Nauss et al."
always_allow_html: true
editor_options:
  chunk_output_type: console
output:
  html_document:
    df_print: paged
  pdf_document: default
---


```{r, echo=FALSE, warning=FALSE, message=FALSE}

# Set up working environment and defaults.
library(envimaR)
if (Sys.info()[["nodename"]] == "PC19616") {
  source("~/plygrnd/CovidAirPolution/CovidAirPolution/src/functions/000_setup.R")
} else {
  source("~/project/cov/CovidAirPolution/src/functions/000_setup.R")
}

start_date <- as.POSIXct("2020-02-15")
end_date <- as.POSIXct("2020-04-20")

pm_vars <- c("PM2.5", "PM10")

lag_vars_set_lut <- (c(
  "median", "median, outlier removed w/o dust event", "median, outlier removed",
  "mean", "mean, outlier removed w/o dust event", "mean, outlier removed"
))
names(lag_vars_set_lut) <- c(
  "pm_median", "pm_median_estm", "pm_median_estm_best",
  "pm_mean", "pm_mean_estm", "pm_mean_estm_best"
)

lag_var <- "pm_median_estm_best"

Sys.setlocale("LC_TIME", "English")

cmpldata <- lapply(pm_vars, function(pm) {
  cmpldata_file <- paste0(pm, "_germany_extended.RDS")
  return(readRDS(file.path(envrmt$path_analysis, cmpldata_file)))
})
names(cmpldata) <- pm_vars

```





# Countrywide temporal development of PM values and SARS-CoV2 infections

This analysis part is based on countrywide averaged data. A GAM model is computed to predict the daily new cases explained by time. The figures show the mean daily PM and covid cases dynamics as well as the daily new cases if time is used as explanatory variable for SARS-CoV2 infections.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_cntry_avg <- lapply(pm_vars, function(pm){
  cntry_avg <- cmpldata[[pm]]$de_avg
  
  cntry_avg_gam <- cntry_avg[cntry_avg$date >= as.POSIXct(start_date) & cntry_avg$date <= as.POSIXct(end_date), ]
  gam_model <- gam(new_cases ~ s(seq(length(date))) + weekday, family = quasipoisson, data = cntry_avg_gam)
  cntry_avg_gam$predicted <- predict(gam_model, cntry_avg_gam, type = "response")
  cntry_avg_gam$residuals <- residuals(gam_model, type = "response")
  
  cntry_indv <- cmpldata[[pm]]$de_clstr$clstr
  
  figure_cntry_avg <- ggplot() +
    geom_line(data = cntry_indv, aes(x = date, y = get(lag_var), group = nuts3Code, color = "All regions")) +
    geom_line(data = cntry_avg_gam, aes(x = date, y = get(lag_var), color = "Daily median PM")) +
    geom_point(data = cntry_avg_gam, aes(x = date, y = get(lag_var), color = "Daily median PM")) +
    geom_line(data = cntry_avg_gam, aes(x = date, y = new_cases, color = "Daily new cases")) +
    geom_point(data = cntry_avg_gam, aes(x = date, y = new_cases, color = "Daily new cases")) +
    geom_line(data = cntry_avg_gam, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
    geom_point(data = cntry_avg_gam, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
    geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype = "dotted", color = "black") +
    geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype = "dotted", color = "black") +
    labs(x = "Date and day of week", y = bquote("Infections, " ~ .(cntry_indv$pm_size[1]) ~ " [" ~ µm / m^3 ~ "]")) +
    theme_bw() +
    scale_x_datetime(date_labels = "%d.%m, %a", date_breaks = "2 day", date_minor_breaks = "1 day") +
    scale_color_manual(
      values = c("#4f4f4f", "#b2df8a", "#1f78b4", "#a6cee3"),
      labels = c(
        paste0("Daily median ", pm), paste0("Daily averaged median ", cntry_indv$pm_size[1]), "Daily new cases",
        "Daily new cases, explained by time"
      )
    ) +
    theme(
      text = element_text(size = 10), axis.title = element_text(size = 10),
      legend.position = c(0.2, 0.7), axis.text.x = element_text(angle = 90, hjust = 0.95, vjust = 0.2),
      legend.title = element_blank(), panel.background = element_blank(),
      panel.grid.major = element_blank(), panel.grid.minor = element_blank()
    )
  return(figure_cntry_avg)
})
names(figure_cntry_avg) <- pm_vars

```





# Correlation of PM and SARS-CoV2 infections individually by district

This analysis part is based on individual rural district data. A GAM model is computed to evaluate the influence of lagged PM concentrations on daily new SARS-CoV2 infections. The figures show the effect size of the model for the individual time lags.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_gam_lag_set <- lapply(pm_vars, function(pm){ 
  
  cntry_indv <- cmpldata[[pm]]$de_clstr$clstr
  gam_lag_set <- readRDS(file.path(envrmt$path_analysis, paste0(pm, "_gam_lag_vars_set.rds")))

  
  # Make figure (set).
  figure_gam_lag_set <- lapply(names(gam_lag_set), function(n) {
    gam_lag <- gam_lag_set[[n]]
    
    tstat <- lapply(seq(0, -15), function(i) {
      tmp <- t.test(gam_lag$t[gam_lag$lag == i])
      data.frame(
        lag = i,
        t = tmp$statistic,
        p = tmp$p.value,
        estimate = tmp$estimate,
        conf_min = tmp$conf.int[1],
        conf_max = tmp$conf.int[2]
      )
    })
    tstat <- do.call("rbind", tstat)
    
    gam_lag$fillcol <- FALSE
    gam_lag$fillcol[gam_lag$lag %in% tstat[tstat$p < 0.1, "lag"]] <- TRUE
    
    figure_gam_effectsize <- ggplot(data = gam_lag, aes(x = as.factor(lag), y = t, fill = fillcol, alpha = fillcol)) +
      geom_boxplot() +
      theme_bw() +
      geom_hline(yintercept = c(0), linetype = "dashed", color = "black") +
      labs(x = "Time lag [days]", y = paste0("Effect size, ", cntry_indv$pm_size[1], " ", lag_vars_set_lut[n])) +
      theme(
        text = element_text(size = 10), axis.title = element_text(size = 10),
        panel.background = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ) +
      scale_fill_manual(values = c("grey", "darkgreen")) +
      scale_alpha_manual(values = c(0.2, 0.6)) +
      theme(legend.position = "none")
    return(figure_gam_effectsize)
  })
  names(figure_gam_lag_set) <- names(gam_lag_set)
  return(figure_gam_lag_set)
})
names(figure_gam_lag_set) <- pm_vars

```





# Correlation of PM and SARS-CoV2 infections with district as random effect

This analysis part is based on individual rural district data. A mixed effect GAM model with district as random effect is computed to evaluate the influence of lagged PM concentrations on daily new SARS-CoV2 infections. The figure shows the effect size of the model for the individual time lags.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_gam_lag_mixed_set <- lapply(pm_vars, function(pm){ 
  
  cntry_indv <- cmpldata[[pm]]$de_clstr$clstr
  gam_lag_mixed_set <- readRDS(file.path(envrmt$path_analysis, paste0(pm, "_gamm_lag_mixed_vars_set.rds")))

  
  # Make figure (set).
  figure_gam_lag_mixed_set <- lapply(names(gam_lag_mixed_set), function(n) {
    gam_lag_mixed <- gam_lag_mixed_set[[n]]
    
    gam_lag_mixed$fillcol <- FALSE
    gam_lag_mixed$fillcol[gam_lag_mixed$p_lme < 0.1] <- TRUE
    
    figure_effectsize_mixed <- ggplot(
      data = gam_lag_mixed,
      aes(x = as.factor(lag), y = t_lme, color = fillcol, fill = fillcol, alpha = fillcol)
    ) +
      geom_point(size = 3) +
      theme_bw() +
      geom_hline(yintercept = c(0), linetype = "dashed", color = "black") +
      labs(x = "Time lag [days]", y = paste0("Effect size, ", cntry_indv$pm_size[1], " ", lag_vars_set_lut[n])) +
      theme(
        text = element_text(size = 10), axis.title = element_text(size = 10),
        panel.background = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ) +
      scale_fill_manual(values = c("black", "red")) +
      scale_color_manual(values = c("black", "red")) +
      scale_alpha_manual(values = c(1, 1)) +
      theme(legend.position = "none")
    figure_effectsize_mixed
    
    figure_estimate_mixed <- ggplot(
      data = gam_lag_mixed,
      aes(x = as.factor(lag), y = pm_lme_estimate, color = fillcol, fill = fillcol, alpha = fillcol)
    ) +
      geom_point() +
      geom_line() +
      geom_errorbar(aes(ymin = pm_lme_estimate - pm_lme_std_error, ymax = pm_lme_estimate + pm_lme_std_error)) +
      theme_bw() +
      geom_hline(yintercept = c(0), linetype = "dashed", color = "black") +
      labs(x = "Time lag [days]", y = paste0("Estimate, ", cntry_indv$pm_size[1], " ", lag_vars_set_lut[n])) +
      theme(
        text = element_text(size = 10), axis.title = element_text(size = 10),
        panel.background = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ) +
      scale_fill_manual(values = c("black", "red")) +
      scale_color_manual(values = c("black", "red")) +
      scale_alpha_manual(values = c(1, 1)) +
      theme(legend.position = "none")
    figure_estimate_mixed
    
    return(list(effect = figure_effectsize_mixed, estimate_mixed = figure_estimate_mixed))
  })
  names(figure_gam_lag_mixed_set) <- names(gam_lag_mixed_set)
  return(figure_gam_lag_mixed_set)
})
names(figure_gam_lag_mixed_set) <- pm_vars

```





# Longer-term correlation of PM and SARS-CoV2 infections

This analysis part is based on individual rural district data that is averaged over time from the start of the data series to April 1st. A linear model is computed to evaluate the influence of longer-term averaged PM concentrations on cummulative SARS-CoV2 infections on April 1st. The figure shows scatterplots between the cumulative infection number and a set of explanatory variables used within the linear model.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

figure_cumulative_effect <- lapply(pm_vars, function(pm){ 
  
  cntry_indv <- cmpldata[[pm]]$de_clstr$clstr
  
  add_info <- lapply(unique(cntry_indv$nuts3Code), function(n) {
    tmp <- cntry_indv[cntry_indv$nuts3Code == n, ]
    tmp$date_numeric <- seq(nrow(tmp))
    
    tmp <- data.frame(
      nuts3Code = tmp$nuts3Code[1],
      date_start = tmp$date[which(tmp$cases > 0)[1]],
      date_max = tmp$date[tail(which(tmp$new_cases == max(tmp$new_cases)), n = 1)],
      daily_max = tmp$new_cases[tail(which(tmp$new_cases == max(tmp$new_cases)), n = 1)],
      cases_at_max = tmp$cases[tail(which(tmp$new_cases == max(tmp$new_cases)), n = 1)],
      cases_on_0401 = tmp$cases[tmp$date == as.POSIXct("2020-04-01")]
    )
    return(tmp)
  })
  
  add_info <- do.call("rbind", add_info)
  add_info$difftime_start_max <- difftime(add_info$date_max, add_info$date_start, units = "days")
  add_info$difftime_first_shutdown <- difftime(add_info$date_start, as.POSIXct("2020-03-17"), units = "days")
  add_info$difftime_max_shutdown <- difftime(add_info$date_max, as.POSIXct("2020-03-17"), units = "days")
  
  tmp <- cntry_indv[cntry_indv$date <= as.POSIXct("2020-04-01"), ]
  
  cntry_agg <- aggregate(st_drop_geometry(tmp[lag_var]), by = list(tmp$nuts3Code), FUN = mean)
  colnames(cntry_agg) <- c("nuts3Code", "PM_tavrg")
  
  tmp <- tmp[tmp$date == as.POSIXct("2020-04-01"), ]
  cntry_agg <- merge(cntry_agg, tmp)
  cntry_agg <- merge(cntry_agg, add_info)
  
  cntry_agg$cases_log10 <- log10(cntry_agg$cases)
  cntry_agg$PM_tavrg_log10 <- log10(cntry_agg$PM_tavrg)
  cntry_agg$pop_total_log10 <- log10(cntry_agg$pop_total)
  cntry_agg$st_area_log10 <- log10(cntry_agg$st_area)
  
  
  # Test of the long-term correlation between PM and SARS-CoV2 infections
  tavrg <- cntry_agg[, c(
    "cases_log10", "centroid_lat", "centroid_lon", "pop_total_log10", "st_area_log10",
    "difftime_first_shutdown", "PM_tavrg_log10"
  )]
  
  lm_tavrg <- lm(cases_log10 ~ PM_tavrg_log10 + centroid_lat + centroid_lon +
                   pop_total_log10 + st_area_log10 + difftime_first_shutdown, data = tavrg)
  summary(lm_tavrg)
  
  lm_tavrg <- lm(cases_log10 ~ 1, data = tavrg)
  lm_tavrg_stpAIC <- stepAIC(lm_tavrg,
                             direction = "both",
                             scope = list(
                               lower = lm_tavrg,
                               upper = ~ PM_tavrg_log10 + centroid_lat + centroid_lon +
                                 pop_total_log10 + st_area_log10 + difftime_first_shutdown
                             )
  )
  summary(lm_tavrg_stpAIC)
  
  tavrg_long <- gather(tavrg, key = "var", value = "value", -cases_log10)
  tavrg_long$var <- factor(tavrg_long$var,
                           levels = c(
                             "PM_tavrg_log10", "centroid_lat", "centroid_lon",
                             "pop_total_log10", "st_area_log10",
                             "difftime_first_shutdown"
                           ),
                           labels = c(
                             paste0("log10(", tmp$pm_size[1], ")"), "Latitude", "Longitude",
                             "log10(Population)", "log10(Area)",
                             "Time difference first case and shutdown"
                           )
  )
  
  figure_cumulative_effect <- ggplot(tavrg_long, aes(x = value, y = cases_log10)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(x = "", y = bquote(log[10] ~ "(Cumulative COVID-19 cases)")) +
    facet_wrap(~var, scales = "free") +
    theme_bw()
  return(figure_cumulative_effect)
})
names(figure_cumulative_effect) <- pm_vars

```





# Maps

Create some maps based on previous results.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

# SARS-CoV2 infections on 2020-04-01
map_covid_infections <- lapply(pm_vars, function(pm){ 
  background_map <- st_transform(ne_countries(scale = "large", continent = "europe", returnclass = "sf"),
                                 crs = "+init=epsg:25832"
  )
  layer <- st_transform(cmpldata[[pm]]$de_nuts3_map, crs = "+init=epsg:25832")
  map_covid_infections <- ggplot(data = background_map) +
    geom_sf() +
    geom_sf(data = layer, aes(fill = cases)) +
    scale_fill_viridis_c(trans = "log10") +
    theme(
      text = element_text(size = 10), axis.title = element_text(size = 10),
      legend.position = "bottom"
    ) +
    labs(fill = "Cumulative infections on 01.04.2020") +
    coord_sf(
      xlim = c(
        (st_bbox(layer)["xmin"] - 10000),
        (st_bbox(layer)["xmax"] + 10000)
      ),
      ylim = c(
        (st_bbox(layer)["ymin"] - 25000),
        (st_bbox(layer)["ymax"] + 10000)
      ), expand = FALSE
    ) +
    annotation_north_arrow(
      location = "tl", which_north = "true",
      height = unit(0.75, "cm"), width = unit(0.75, "cm"),
      pad_x = unit(0.0, "in"), pad_y = unit(0.1, "in"),
      style = north_arrow_minimal
    )
  return(map_covid_infections)
})
names(map_covid_infections) <- pm_vars


# Long-term PM median between 2020-02-15 and 2020-04-01
map_pm_median <- lapply(pm_vars, function(pm){ 
  background_map <- st_transform(ne_countries(scale = "large", continent = "europe", returnclass = "sf"),
                                 crs = "+init=epsg:25832"
  )
  layer <- st_transform(cmpldata[[pm]]$de_nuts3_map, crs = "+init=epsg:25832")
  map_pm_median <- ggplot(data = background_map) +
    geom_sf() +
    geom_sf(data = layer, aes(fill = pm_median)) +
    scale_fill_viridis_c(trans = "log10") +
    theme(
      text = element_text(size = 10), axis.title = element_text(size = 10),
      legend.position = "bottom"
    ) +
    labs(fill = paste0("Median ", cmpldata[[pm]]$de_nuts3_map$pm_size[1], " 15.02.2020 to 01.04.2020")) +
    coord_sf(
      xlim = c(
        (st_bbox(layer)["xmin"] - 10000),
        (st_bbox(layer)["xmax"] + 10000)
      ),
      ylim = c(
        (st_bbox(layer)["ymin"] - 25000),
        (st_bbox(layer)["ymax"] + 10000)
      ), expand = FALSE
    ) +
    annotation_north_arrow(
      location = "tl", which_north = "true",
      height = unit(0.75, "cm"), width = unit(0.75, "cm"),
      pad_x = unit(0.0, "in"), pad_y = unit(0.1, "in"),
      style = north_arrow_minimal
    )
  return(map_pm_median)
})
names(map_pm_median) <- pm_vars
```





# Figures for publication
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
figure_cntry_avg <- readRDS(file.path(envrmt$path_figures, "figure_cntry_avg.rds"))
figure_gam_lag_set <- readRDS(file.path(envrmt$path_figures, "figure_gam_lag_set.rds"))
figure_gam_lag_mixed_set <- readRDS(file.path(envrmt$path_figures, "figure_gam_lag_mixed_set.rds"))
figure_cumulative_effect <- readRDS(file.path(envrmt$path_figures, "figure_cumulative_effect.rds"))
map_covid_infections <- readRDS(file.path(envrmt$path_figures, "map_covid_infections.rds"))
map_pm_median <- readRDS(file.path(envrmt$path_figures, "map_pm_median.rds"))
lm_tavrg_smry <- readRDS(file.path(envrmt$path_figures, "lm_tavrg_smry.rds"))
lm_tavrg_stpAIC_smry <- readRDS(file.path(envrmt$path_figures, "lm_tavrg_stpAIC_smry.rds"))

# names(.GlobalEnv)[grep("figure", names(.GlobalEnv))]
# names(.GlobalEnv)[grep("map", names(.GlobalEnv))]
# plot(rd)


figure_1_1 <- plot_grid(figure_cntry_avg, figure_cntry_avg_res,
                        labels = c("a", "b"), label_size = 10, ncol = 1
)
figure_1_2 <- plot_grid(map_covid_infections, map_pm_median,
                        labels = c("c", "d"), label_size = 10, ncol = 2
)
figure_1 <- plot_grid(figure_1_1, figure_1_2,
                      labels = NULL, ncol = 1
)

ggsave(file.path(envrmt$path_figures, "figure_1.png"),
       plot = figure_1,
       width = 216, height = 279, units = "mm", dpi = 300
)



pdf(NULL, width = 5, height = 10)
dev.control(displaylist = "enable")
plot(rd, type = "points", scaling = 0)
figure_rda <- recordPlot()
invisible(dev.off())

figure_2_1 <- plot_grid(figure_cumulative_effect,
                        labels = c("a"), label_size = 10, ncol = 1
)

figure_2_2 <- plot_grid(figure_effectsize, figure_estimate_mixed,
                        labels = c("b", "c"), label_size = 10, ncol = 2
)

figure_2_2_alternative <- plot_grid(figure_effectsize, figure_effectsize_mixed,
                                    labels = c("b", "c"), label_size = 10, ncol = 2
)

figure_2 <- plot_grid(figure_2_1, figure_2_2,
                      labels = NULL, ncol = 1
)

figure_2_alternative <- plot_grid(figure_2_1, figure_2_2_alternative,
                                  labels = NULL, ncol = 1
)

ggsave(file.path(envrmt$path_figures, "figure_2.png"),
       plot = figure_2,
       width = 216, height = 279, units = "mm", dpi = 300
)

ggsave(file.path(envrmt$path_figures, "figure_2_alternative.png"),
       plot = figure_2_alternative,
       width = 216, height = 279, units = "mm", dpi = 300
)
```


















































# Geographical distribution of SARS-CoV2 infections in cmpldata
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = TRUE}
print(paste("Start date for analysis: ", start_date))
print(paste("End date for analysis:   ", end_date))
print(paste("PM values:               ", pm))

cmpldata$de_nuts3_map$log_cases <- log10(cmpldata$de_nuts3_map$cases)
mapview(cmpldata$de_nuts3_map, zcol = "log_cases")




figure_cntry_avg_res <- ggplot() +
  geom_line(data = cntry_avg_gam, aes(x = date, y = residuals, color = "Residuals")) +
  geom_point(data = cntry_avg_gam, aes(x = date, y = residuals, color = "Residuals")) +
  geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype = "dotted", color = "black") +
  geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype = "dotted", color = "black") +
  geom_text(aes(
    x = as.POSIXct("2020-03-14"),
    label = "Start of contact restrictions \n", y = -0.6
  ), colour = "black", angle = 90, size = 2) +
  geom_text(aes(
    x = as.POSIXct("2020-03-17"),
    label = "Start of shut down \n", y = -0.65
  ), colour = "black", angle = 90, size = 2) +
  labs(x = "Date and day of week", y = "Residuals") +
  theme_bw() +
  scale_x_datetime(date_labels = "%d.%m, %a", date_breaks = "2 day", date_minor_breaks = "1 day") +
  scale_color_manual(values = c("#d95f02")) +
  theme(
    text = element_text(size = 10), axis.title = element_text(size = 10),
    legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 0.95, vjust = 0.2),
    legend.title = element_blank(), panel.background = element_blank(),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  )
figure_cntry_avg_res
```







# Redundancy analysis of the lagged effect sizes

The analysis part is based on former and additional results and aims in providing additional information regarding the dependencies of SARS-CoV2 infections. The figure shows the results of a redundancy analysis of the lagged effect sizes.

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

# summary(as.numeric(add_info$difftime_start_max))
# summary(as.numeric(add_info$first_shutdown))
# summary(as.numeric(add_info$max_shutdown))
# summary(add_info$daily_max)
#
# hist(as.numeric(add_info$difftime_start_max))
# hist(as.numeric(add_info$first_shutdown))
# hist(as.numeric(add_info$max_shutdown))

t_matrix <- reshape(gam_lag[, c("nuts3Code", "t", "lag")], idvar = "nuts3Code", timevar = "lag", direction = "wide")
rownames(t_matrix) <- t_matrix$nuts3Code

t_environment <- merge(cmpldata$de_nuts3_map, add_info)
t_environment <- st_drop_geometry(t_environment)
t_matrix <- merge(t_matrix, t_environment, by = "nuts3Code")

# Log transform some variables and rename for figure.
t_matrix$PM <- log10(t_matrix$pm_median)
t_matrix$Population <- log10(t_matrix$pop_total)
t_matrix$Area <- log10(t_matrix$st_area)
t_matrix$Cases <- log10(t_matrix$cases)
t_matrix$First <- t_matrix$difftime_first_shutdown
t_matrix$Latitude <- t_matrix$centroid_lat
t_matrix$Longitude <- t_matrix$centroid_lon

rd <- rda(
  t_matrix[, seq(which(colnames(t_matrix) == "t.0"), which(colnames(t_matrix) == "t.-15"))],
  t_matrix[, c("PM", "Cases", "Population", "Area", "First", "Latitude", "Longitude")]
)
plot(rd, type = "points", scaling = 0)
```







# Maps
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
# SARS-CoV2 infections on 2020-04-01
background_map <- st_transform(ne_countries(scale = "large", continent = "europe", returnclass = "sf"),
                               crs = "+init=epsg:25832"
)
layer <- st_transform(cmpldata$de_nuts3_map, crs = "+init=epsg:25832")
map_covid_infections <- ggplot(data = background_map) +
  geom_sf() +
  geom_sf(data = layer, aes(fill = cases)) +
  scale_fill_viridis_c(trans = "log10") +
  theme(
    text = element_text(size = 10), axis.title = element_text(size = 10),
    legend.position = "bottom"
  ) +
  labs(fill = "Infections (01.04.2020)") +
  coord_sf(
    xlim = c(
      (st_bbox(layer)["xmin"] - 10000),
      (st_bbox(layer)["xmax"] + 10000)
    ),
    ylim = c(
      (st_bbox(layer)["ymin"] - 25000),
      (st_bbox(layer)["ymax"] + 10000)
    ), expand = FALSE
  ) +
  annotation_north_arrow(
    location = "tl", which_north = "true",
    height = unit(0.75, "cm"), width = unit(0.75, "cm"),
    pad_x = unit(0.0, "in"), pad_y = unit(0.1, "in"),
    style = north_arrow_minimal
  )
map_covid_infections


# Long-term PM median between 2020-02-15 and 2020-04-01
background_map <- st_transform(ne_countries(scale = "large", continent = "europe", returnclass = "sf"),
                               crs = "+init=epsg:25832"
)
layer <- st_transform(cmpldata$de_nuts3_map, crs = "+init=epsg:25832")
map_pm_median <- ggplot(data = background_map) +
  geom_sf() +
  geom_sf(data = layer, aes(fill = pm_median)) +
  scale_fill_viridis_c(trans = "log10") +
  theme(
    text = element_text(size = 10), axis.title = element_text(size = 10),
    legend.position = "bottom"
  ) +
  labs(fill = "PM10 median 15.02.2020 - 01.04.2020") +
  coord_sf(
    xlim = c(
      (st_bbox(layer)["xmin"] - 10000),
      (st_bbox(layer)["xmax"] + 10000)
    ),
    ylim = c(
      (st_bbox(layer)["ymin"] - 25000),
      (st_bbox(layer)["ymax"] + 10000)
    ), expand = FALSE
  ) +
  annotation_north_arrow(
    location = "tl", which_north = "true",
    height = unit(0.75, "cm"), width = unit(0.75, "cm"),
    pad_x = unit(0.0, "in"), pad_y = unit(0.1, "in"),
    style = north_arrow_minimal
  )
map_pm_median



# RDA1 values
score_geo <- cbind(t_matrix$nuts3Code, scores(rd)$sites)
colnames(score_geo)[1] <- "nuts3Code"
score_geo <- merge(score_geo, unique(de_clstr[, "nuts3Code"]))
st_geometry(score_geo) <- score_geo$geometry

background_map <- st_transform(ne_countries(scale = "large", continent = "europe", returnclass = "sf"),
                               crs = "+init=epsg:25832"
)
layer <- st_transform(score_geo, crs = "+init=epsg:25832")
map_rda1_score <- ggplot(data = background_map) +
  geom_sf() +
  geom_sf(data = layer, aes(fill = RDA1)) +
  theme(legend.position = "bottom") +
  labs(fill = "RDA1") +
  coord_sf(
    xlim = c(
      (st_bbox(layer)["xmin"] - 10000),
      (st_bbox(layer)["xmax"] + 10000)
    ),
    ylim = c(
      (st_bbox(layer)["ymin"] - 25000),
      (st_bbox(layer)["ymax"] + 10000)
    ), expand = FALSE
  ) +
  annotation_north_arrow(
    location = "tl", which_north = "true",
    height = unit(0.75, "cm"), width = unit(0.75, "cm"),
    pad_x = unit(0.0, "in"), pad_y = unit(0.1, "in"),
    style = north_arrow_minimal
  )
map_rda1_score



# Significant and positive t values
p <- 0.1

lag_analysis <- st_set_geometry(
  gam_lag_geo$test_geo[gam_lag_geo$test_geo$t > 0 &
                         gam_lag_geo$test_geo$p < p, ],
  gam_lag_geo$test_geo$geometry[gam_lag_geo$test_geo$t > 0 &
                                  gam_lag_geo$test_geo$p < p, ]
)

lag_analysis <- lapply(unique(lag_analysis$nuts3Code), function(n) {
  data.frame(
    nuts3Code = lag_analysis$nuts3Code[lag_analysis$nuts3Code == n][1],
    min_lag = max(lag_analysis$lag[lag_analysis$nuts3Code == n]),
    mean_lag = mean(lag_analysis$lag[lag_analysis$nuts3Code == n]),
    max_lag = min(lag_analysis$lag[lag_analysis$nuts3Code == n]),
    geometry = lag_analysis$geometry[lag_analysis$nuts3Code == n][1]
  )
})
lag_analysis <- do.call("rbind", lag_analysis)
lag_analysis <- merge(lag_analysis, cmpldata$de_nuts3_map[, c("nuts3Code", "pm_median", "cases", "pop_dens")])
hist(lag_analysis$min_lag)
hist(lag_analysis$mean_lag)
hist(lag_analysis$max_lag)

non_significant_regions <- cmpldata$de_nuts3_map[!(cmpldata$de_nuts3_map$nuts3Code %in% lag_analysis$nuts3Code), ]

summary(cmpldata$de_nuts3_map[, c("nuts3Code", "pm_median", "cases", "pop_dens")])
summary(lag_analysis)
summary(non_significant_regions[, c("nuts3Code", "pm_median", "cases", "pop_dens")])

quantile(non_significant_regions$pop_dens, probs = seq(0, 1, 0.1))
quantile(lag_analysis$pop_dens, probs = seq(0, 1, 0.1))

wilcox.test(non_significant_regions$cases, lag_analysis$cases)
wilcox.test(non_significant_regions$pm_median, lag_analysis$pm_median)
wilcox.test(non_significant_regions$pop_dens, lag_analysis$pop_dens)

t.test(non_significant_regions$cases, lag_analysis$cases)
t.test(non_significant_regions$pm_median, lag_analysis$pm_median)
t.test(non_significant_regions$pop_dens, lag_analysis$pop_dens)

layer1 <- st_set_geometry(lag_analysis, lag_analysis$geometry)
layer1 <- st_transform(layer1, crs = "+init=epsg:25832")

background_map <- st_transform(ne_countries(scale = "large", continent = "europe", returnclass = "sf"),
                               crs = "+init=epsg:25832"
)
layer <- st_transform(cmpldata$de_nuts3_map, crs = "+init=epsg:25832")

map_t_sig <- ggplot(data = background_map) +
  geom_sf() +
  geom_sf(data = layer) +
  geom_sf(data = layer1, aes(fill = mean_lag)) +
  scale_fill_viridis_c() +
  theme(legend.position = "bottom") +
  labs(fill = "Mean lag of positive effect size") +
  coord_sf(
    xlim = c(
      (st_bbox(layer)["xmin"] - 10000),
      (st_bbox(layer)["xmax"] + 10000)
    ),
    ylim = c(
      (st_bbox(layer)["ymin"] - 25000),
      (st_bbox(layer)["ymax"] + 10000)
    ), expand = FALSE
  ) +
  annotation_north_arrow(
    location = "tl", which_north = "true",
    height = unit(0.5, "cm"), width = unit(0.5, "cm"),
    pad_x = unit(0.0, "in"), pad_y = unit(0.1, "in"),
    style = north_arrow_minimal
  )
map_t_sig
```


# Figures for publication
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}


figure_effectsize
plot(rd)
map_covid_infections
map_pm_median


figure_1_1 <- plot_grid(figure_cntry_avg, figure_cntry_avg_res,
                        labels = c("a", "b"), label_size = 10, ncol = 1
)
figure_1_2 <- plot_grid(map_covid_infections, map_pm_median,
                        labels = c("c", "d"), label_size = 10, ncol = 2
)
figure_1 <- plot_grid(figure_1_1, figure_1_2,
                      labels = NULL, ncol = 1
)

# ggsave(file.path(envrmt$path_figures, "figure_cntry_avg.png"), plot = figure_cntry_avg,
#        width = 216, height = 279, units = "mm", dpi = 300)
# ggsave(file.path(envrmt$path_figures, "figure_cntry_avg_res.png"), plot = figure_cntry_avg_res,
#        width = 216, height = 279, units = "mm", dpi = 300)
# ggsave(file.path(envrmt$path_figures, "map_covid_infections.png"), plot = map_covid_infections,
#        width = 279, height = 216, units = "mm", dpi = 300)
# ggsave(file.path(envrmt$path_figures, "map_pm_median.png"), plot = map_pm_median,
#        width = 279, height = 216, units = "mm", dpi = 300)

ggsave(file.path(envrmt$path_figures, "figure_1.png"),
       plot = figure_1,
       width = 216, height = 279, units = "mm", dpi = 300
)



pdf(NULL, width = 5, height = 10)
dev.control(displaylist = "enable")
plot(rd, type = "points", scaling = 0)
figure_rda <- recordPlot()
invisible(dev.off())

figure_2_1 <- plot_grid(figure_effectsize, figure_rda,
                        labels = c("a", "b"), label_size = 10,
                        rel_widths = c(1, 2)
)

figure_2_2 <- plot_grid(map_rda1_score, map_t_sig,
                        labels = c("c", "d"), label_size = 10, ncol = 2
)

figure_2 <- plot_grid(figure_2_1, figure_2_2,
                      labels = NULL, ncol = 1
)


# ggsave(file.path(envrmt$path_figures, "figure_effectsize.png"), plot = figure_effectsize,
#        width = 216, height = 279, units = "mm", dpi = 300)
# figure_rda_tmp = plot_grid(figure_rda, labels = "")
#
# ggsave(file.path(envrmt$path_figures, "figure_rda.png"), plot = figure_rda_tmp,
#        width = 279, height = 216, units = "mm", dpi = 300)
# ggsave(file.path(envrmt$path_figures, "map_rda1_score.png"), plot = map_rda1_score,
#        width = 279, height = 216, units = "mm", dpi = 300)
# ggsave(file.path(envrmt$path_figures, "map_t_sig.png"), plot = map_t_sig,
#        width = 279, height = 216, units = "mm", dpi = 300)


ggsave(file.path(envrmt$path_figures, "figure_2.png"),
       plot = figure_2,
       width = 216, height = 279, units = "mm", dpi = 300
)
```






# Dynmiac time warp clustering on countrywide average
```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = TRUE}
mapview(cmpldata$de_clstr$clstr_map, zcol = "cluster_covid")
mapview(cmpldata$de_clstr$clstr_map, zcol = "cluster_pm")
```

```{r, eval = TRUE, echo=FALSE, warning=FALSE, message=FALSE}
de_clstr <- cmpldata$de_clstr

clstr_avg_pm_gam_time <- de_clstr$clstr_avg_pm[
  de_clstr$clstr_avg_pm$date >= as.POSIXct("2020-02-15") &
    de_clstr$clstr_avg_pm$date <= as.POSIXct("2020-04-01"),
  ]
clstr_avg_pm_gam_time$weekday <- weekdays(clstr_avg_pm_gam_time$date)
clstr_avg_pm_gam_time$predicted <- NA
clstr_avg_pm_gam_time$residuals <- NA


for (c in unique(clstr_avg_pm_gam_time$cluster_pm)) {
  set.seed(01042020)
  gam_time <- gam(new_cases ~ s(seq(length(date))) + weekday,
                  family = quasipoisson,
                  data = clstr_avg_pm_gam_time[clstr_avg_pm_gam_time$cluster_pm == c, ]
  )
  clstr_avg_pm_gam_time[clstr_avg_pm_gam_time$cluster_pm == c, "predicted"] <- predict(gam_time, type = "response")
  clstr_avg_pm_gam_time[clstr_avg_pm_gam_time$cluster_pm == c, "residuals"] <- residuals(gam_time, type = "response")
}

clstr_avg_covid_gam_time <- de_clstr$clstr_avg_covid[
  de_clstr$clstr_avg_covid$date >= as.POSIXct("2020-02-15") &
    de_clstr$clstr_avg_covid$date <= as.POSIXct("2020-04-01"),
  ]
clstr_avg_covid_gam_time$weekday <- weekdays(clstr_avg_covid_gam_time$date)
clstr_avg_covid_gam_time$predicted <- NA
clstr_avg_covid_gam_time$residuals <- NA

for (c in unique(clstr_avg_covid_gam_time$cluster_covid)) {
  set.seed(01042020)
  gam_time <- gam(new_cases ~ s(seq(length(date))) + weekday,
                  family = quasipoisson,
                  data = clstr_avg_covid_gam_time[clstr_avg_covid_gam_time$cluster_covid == c, ]
  )
  clstr_avg_covid_gam_time[clstr_avg_covid_gam_time$cluster_covid == c, "predicted"] <- predict(gam_time, type = "response")
  clstr_avg_covid_gam_time[clstr_avg_covid_gam_time$cluster_covid == c, "residuals"] <- residuals(gam_time, type = "response")
}

ggplot() +
  geom_line(data = clstr_avg_pm_gam_time, aes(x = date, y = pm_median, color = "Daily median PM10")) +
  geom_line(data = clstr_avg_pm_gam_time, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_line(data = clstr_avg_pm_gam_time, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
  geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype = "dotted", color = "black") +
  geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype = "dotted", color = "black") +
  labs(x = "Date and day of week", y = bquote("Infections, PM10 " ~ µm / m^3)) +
  theme_bw() +
  scale_x_datetime(date_labels = "%d.%m, %a", date_breaks = "5 day", date_minor_breaks = "1 day") +
  scale_color_manual(values = c("#b2df8a", "#1f78b4", "#a6cee3")) +
  theme(
    text = element_text(size = 10), axis.title = element_text(size = 10),
    axis.text.x = element_text(angle = 90, hjust = 0.95, vjust = 0.2),
    legend.title = element_blank(), panel.background = element_blank(),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  ) +
  facet_wrap(~cluster_pm, ncol = 2, nrow = 6, scales = "fix")

ggplot() +
  geom_line(data = clstr_avg_covid_gam_time, aes(x = date, y = pm_median, color = "Daily median PM10")) +
  geom_line(data = clstr_avg_covid_gam_time, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_line(data = clstr_avg_covid_gam_time, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
  geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype = "dotted", color = "black") +
  geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype = "dotted", color = "black") +
  labs(x = "Date and day of week", y = bquote("Infections, PM10 " ~ µm / m^3)) +
  theme_bw() +
  scale_x_datetime(date_labels = "%d.%m, %a", date_breaks = "5 day", date_minor_breaks = "1 day") +
  scale_color_manual(values = c("#b2df8a", "#1f78b4", "#a6cee3")) +
  theme(
    text = element_text(size = 10), axis.title = element_text(size = 10),
    axis.text.x = element_text(angle = 90, hjust = 0.95, vjust = 0.2),
    legend.title = element_blank(), panel.background = element_blank(),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  ) +
  facet_wrap(~cluster_covid, ncol = 2, nrow = 6, scales = "fix")
```



```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
de_clstr <- cmpldata$de_clstr$clstr
# pdata = compilePopulation()
# de_clstr = merge(de_clstr, pdata, by = "nuts3Code")
#
# de_clstr = merge(de_clstr, st_set_geometry(cmpldata$de_nuts3_map[, c("nuts3Code", "nuts3Name", "centroid_lat", "centroid_lon", "nuts3Area")], NULL), by.x = "nuts3Code", by.y = "nuts3Code")
#
# de_clstr$nuts3Area = as.numeric(de_clstr$nuts3Area)
# de_clstr$cases_per_area = de_clstr$cases / de_clstr$nuts3Area
# de_clstr$cases_per_pop = de_clstr$cases / de_clstr$pop_total
# de_clstr$cases_per_pop_dens = de_clstr$cases / de_clstr$pop_dens
# de_clstr$new_cases_per_area = de_clstr$new_cases / de_clstr$nuts3Area
# de_clstr$new_cases_per_pop = de_clstr$new_cases / de_clstr$pop_total
# de_clstr$new_cases_per_pop_dens = de_clstr$new_cases / de_clstr$pop_dens
# de_clstr$pm_mean_dt = c(0, diff(de_clstr$pm_mean))
# de_clstr$weekday_cn = as.numeric(de_clstr$weekday_c)
# de_clstr$days_before_shutdown = difftime(de_clstr$date, as.POSIXct("2020-03-17"), units = "days")


info <- lapply(unique(de_clstr$nuts3Code), function(n) {
  tmp <- de_clstr[de_clstr$nuts3Code == n, ]
  tmp$date_numeric <- seq(nrow(tmp))
  
  tmp <- data.frame(
    nuts = tmp$nuts3Code[1],
    date_start = tmp$date[which(tmp$cases > 0)[1]],
    date_max = tmp$date[tail(which(tmp$new_cases == max(tmp$new_cases)), n = 1)],
    daily_max = tmp$new_cases[tail(which(tmp$new_cases == max(tmp$new_cases)), n = 1)],
    cases_at_max = tmp$cases[tail(which(tmp$new_cases == max(tmp$new_cases)), n = 1)]
  )
  
  # set.seed(01042020)
  # # glm_mod = glm(cases ~ date_numeric + I(date_numeric^2),
  # #               family = gaussian,
  # #               data = tmp[tmp$weekday_c == "W" & tmp$date >= date_start,])
  # glm_mod <- zeroinfl(new_cases ~ date_numeric + weekday_c + pm_mean | days_before_shutdown, data = tmp)
  # tmp$cases_pred = 0
  # tmp$cases_pred[tmp$date >= date_start] = predict(glm_mod, tmp[tmp$date >= date_start,])
  # tmp$cases_pred[tmp$cases_pred < 0] = 0
  # plot(tmp$cases_pred)
  # points(tmp$new_cases, col = "red")
  
  return(tmp)
})
info <- do.call("rbind", info)
info$difftime <- difftime(info$date_max, info$date_start, units = "days")
ggplot(data = info, aes(x = difftime)) +
  geom_histogram(binwidth = 5)
summary(as.numeric(info$difftime[info$cases_at_max > 100]))
summary(info$cases_at_max)
summary(info$daily_max)
summary(info$cases_at_max[info$difftime >= 30])
summary(info$cases_at_max)
summary(as.numeric(info$difftime))

head(de_clstr)

ggplot(data = de_clstr[de_clstr$weekday_c == "M", ], aes(x = date, y = new_cases, color = as.factor(nuts3Code))) +
  geom_line() +
  theme(legend.position = "none")

ndays <- c(-14, 35)
ndays <- c(-14, 20)
ndays <- c(0, 30)

gam_lag <- compileLaggedGLM(
  data = de_clstr[de_clstr$nuts3Code %in% info$nuts[info$difftime >= 30 & info$difftime <= 99], ],
  pm = "org",
  frml = "new_cases ~ date + weekday_c + pm_mean_lag",
  nlags = 20,
  subset_var = "new_cases", subset_thv = 1, individual = "start", ndays = c(0, 35)
)

gam_lag <- compileLaggedGAM(
  data = de_clstr,
  pm = "org",
  frml = "new_cases ~ seq(length(date)) + weekday_c + pm_mean_lag",
  nlags = 20,
  subset_var = "new_cases", subset_thv = 1, individual = "start", ndays = c(0, 35)
)



gam_lag <- compileLaggedGLM(
  data = de_clstr,
  pm = "org",
  frml = "new_cases ~ date + weekday_c + pm_mean_lag",
  nlags = 20,
  subset_var = "new_cases", subset_thv = 1, individual = "start", ndays = c(0, 35)
)



gam_lag_geo <- compileLaggedGLMGeo(gam_lag, cmpldata$de_nuts3_map)

t.test(gam_lag$t[gam_lag$lag == -3])

# gam_lag = gam_lag_org
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
ggplot(data = gam_lag[abs(gam_lag$t) < 5, ], aes(x = as.factor(lag), y = t)) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = "point", shape = 20, size = 5, color = "red", fill = "red") +
  theme_bw() +
  geom_hline(yintercept = c(-2, 2), linetype = "dashed", color = "red") +
  labs(title = paste("Dependence of individual COVID-19 time series and lagged ", pm, " concentrations"), x = "Time lag PM [days]", y = "t value of the partial coefficient")
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
ggplot(data = gam_lag[abs(gam_lag$t) < 5, ], aes(x = as.factor(lag), y = t, fill = cluster_pm)) +
  geom_boxplot() +
  theme_bw() +
  geom_hline(yintercept = c(-2, 2), linetype = "dashed", color = "red") +
  labs(title = paste("Dependence of individual COVID-19 time series and lagged ", pm, " concentrations by PM clusters"), x = "Time lag PM [days]", y = "t value of the partial coefficient")
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
ggplot(data = gam_lag_geo$test_geo_lag) +
  geom_line(aes(x = lag, y = rsq, color = "R squared")) +
  geom_point(aes(x = lag, y = rsq, color = "R squared")) +
  theme_bw() +
  scale_x_continuous(breaks = seq(-15, 0, 1), minor_breaks = seq(-15, 0, 1)) +
  labs(title = paste("R squared values of lagged ", pm, " concentrations based on latitude, longitude and area", x = "Time lag PM [days]"), y = "R squared", color = "R squared")

ggplot(data = gam_lag_geo$test_geo_lag) +
  geom_line(aes(x = lag, y = centroid_lat_t, color = "latitude")) +
  geom_point(aes(x = lag, y = centroid_lat_t, color = "latitude")) +
  geom_line(aes(x = lag, y = centroid_lon_t, color = "longitude")) +
  geom_point(aes(x = lag, y = centroid_lon_t, color = "longitude")) +
  geom_line(aes(x = lag, y = area_t, color = "log10(area)")) +
  geom_point(aes(x = lag, y = area_t, color = "log10(area)")) +
  theme_bw() +
  geom_hline(yintercept = c(-2, 2), linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(-15, 0, 1), minor_breaks = seq(-15, 0, 1)) +
  scale_y_continuous(breaks = seq(-10, 20, 2)) +
  labs(title = paste("t values of lagged ", pm, " concentrations for latitude, longitude and area"), x = "Time lag PM [days]", y = "t value of the partial coefficient", color = "Partial coefficient")

ggplot(data = gam_lag_geo$test_geo[gam_lag_geo$test_geo$lag == -9, ]) +
  geom_point(aes(x = centroid_lat, y = t, color = as.factor(lag))) +
  theme_bw() +
  geom_hline(yintercept = c(-2, 2), linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(47, 55, 1)) +
  scale_y_continuous(breaks = seq(-10, 20, 2)) +
  labs(title = paste("t values of lagged ", pm, " concentrations for latitude"), x = "Latitude [degrees]", y = "t value of the partial coefficient for PM (glm model)", color = "Lag of PM")

ggplot(data = gam_lag_geo$test_geo[gam_lag_geo$test_geo$lag >= -7 & gam_lag_geo$test_geo$lag <= -4, ]) +
  geom_point(aes(x = centroid_lat, y = t, color = as.factor(lag))) +
  theme_bw() +
  geom_hline(yintercept = c(-2, 2), linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(47, 55, 1)) +
  scale_y_continuous(breaks = seq(-10, 20, 2)) +
  labs(title = paste("t values of lagged ", pm, " concentrations for latitude"), x = "Latitude [degrees]", y = "t value of the partial coefficient for PM (glm model)", color = "Lag of PM")
```

Hier mal ein Versuch, die "Italien-Grafik" aus dem  etwas nachzubauen. Die Signifikanz ist bei PM10 für längere Zeitintervalle gegeben, für PM2.5 praktisch nicht (Setti, L. et al. The Potential role of Particulate Matter in the Spreading of COVID-19 in Northern Italy: First Evidence-based Research Hypotheses. medRxiv, https://www.medrxiv.org/content/10.1101/2020.04.11.20061713v1 (2020).)

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
# 10, 12
test <- cmpldata$de_clstr$clstr[cmpldata$de_clstr$clstr$date < as.POSIXct("2020-03-17"), ] %>%
  group_by(nuts3Code) %>%
  summarize(
    n_days = n(),
    pm_gt_10 = sum(pm_mean > 10),
    pm_gt_10_rel = pm_gt_10 / n_days
  )
test_merge <- merge(test, cmpldata$de_clstr$clstr[cmpldata$de_clstr$clstr$date == as.POSIXct("2020-04-01"), ])
test_merge <- merge(test_merge, st_set_geometry(cmpldata$de_nuts3_map, NULL)[, c("nuts3Code", "area")])
test_merge$ratio <- as.numeric(test_merge$cases / test_merge$area)

ggplot(data = test_merge, aes(x = pm_gt_10, y = ratio)) +
  geom_point() +
  geom_smooth(method = "glm") +
  theme_bw() +
  scale_y_log10() +
  labs(title = paste("COVID-19 infections per province area (01.04.) and days exceeding", pm, " > 10 from 15.2. to 17.03."))
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = TRUE}
model <- glm(test_merge$ratio ~ test_merge$pm_gt_10, family = quasipoisson)
summary(model)
```



```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
# Wavelet coherence analysis

## Wavelet coherence analysis on countrywide average
cntry_avg <- cmpldata$cntry_avg
wc <- compileWC(cntry_avg)

wc.image(wc,
         which.image = "wc",
         n.levels = 250, color.key = "interval",
         siglvl.contour = 0.1, siglvl.arrow = 0.1,
         legend.params = list(lab = "cross-wavelet power levels"), show.date = TRUE,
         spec.period.axis = list(at = seq(15)),
         main = paste("Wavelet coherence cmpldata, ", pm, " x detrended daily COVID-19 cases")
)

# wc.sel.phases(wc, sel.period = 4, siglvl = 1, show.date = TRUE)
```



```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
## Wavelet coherence analysis on cluster averages
de_clstr <- cmpldata$de_clstr
for (c in unique(de_clstr$clstr_avg_detr$cluster_pm)) {
  tmp <- de_clstr$clstr_avg_detr[de_clstr$clstr_avg_detr$cluster_pm == c, ]
  wc <- compileWC(tmp)
  
  wc.image(wc,
           which.image = "wc",
           n.levels = 250, color.key = "interval",
           siglvl.contour = 0.1, siglvl.arrow = 0.1,
           legend.params = list(lab = "cross-wavelet power levels"), show.date = TRUE,
           spec.period.axis = list(at = seq(15)),
           main = paste("cmpldata, ", pm, " x detrended daily COVID-19 cases, cluster ", c)
  )
  
  # wc.sel.phases(wc, sel.period = 4, siglvl = 1, show.date = TRUE)
}
```

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}
cntry_avg_glm_time_03 <- cntry_avg[cntry_avg$date >= as.POSIXct("2020-03-01") & cntry_avg$date <= as.POSIXct("2020-04-20"), ]
set.seed(01042020)
cntry_avg_glm_time <- gam(new_cases ~ s(seq(length(date))) + weekday, family = quasipoisson, data = cntry_avg_glm_time_03)
cntry_avg_glm_time <- glm(new_cases ~ seq(length(date)) + weekday, family = quasipoisson, data = cntry_avg_glm_time_03)
cntry_avg_glm_time_03$predicted <- predict(cntry_avg_glm_time, cntry_avg_glm_time_03, type = "response")
cntry_avg_glm_time_03$residuals <- residuals(cntry_avg_glm_time, type = "response")

figure_01 <- ggplot() +
  geom_line(data = cntry_avg, aes(x = date, y = pm_mean, color = "Daily mean PM10")) +
  geom_point(data = cntry_avg, aes(x = date, y = pm_mean, color = "Daily mean PM10")) +
  geom_line(data = cntry_avg, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_point(data = cntry_avg, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_line(data = cntry_avg_glm_time_03, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
  geom_point(data = cntry_avg_glm_time_03, aes(x = date, y = predicted, color = "Daily new cases, explained by time")) +
  geom_line(data = cntry_avg_glm_time_03, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_point(data = cntry_avg_glm_time_03, aes(x = date, y = new_cases, color = "Daily new cases")) +
  geom_vline(xintercept = as.POSIXct("2020-03-14"), linetype = "dotted", color = "black") +
  geom_vline(xintercept = as.POSIXct("2020-03-17"), linetype = "dotted", color = "black") +
  # geom_text(aes(x = as.POSIXct("2020-03-14"), label="Start of contact restrictions \n", y=-1), colour="black", angle=90, size = 3) +
  # geom_text(aes(x = as.POSIXct("2020-03-17"), label="Start of shut down \n", y=-1), colour="black", angle=90, size = 3) +
  labs(x = "Date and day of week", y = expression("Infections, PM10 ~µm/m^{3}")) +
  theme_bw() +
  scale_x_datetime(date_labels = "%d.%m, %a", date_breaks = "2 day", date_minor_breaks = "1 day") +
  scale_color_manual(values = c("#b2df8a", "#1f78b4", "#a6cee3")) +
  theme(
    text = element_text(size = 10), axis.title = element_text(size = 10),
    legend.position = c(0.35, 0.75), axis.text.x = element_text(angle = 90, hjust = 0.95, vjust = 0.2),
    legend.title = element_blank(), panel.background = element_blank(),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  )
figure_01
```




















# Deprecated section from former analysis versions.

## Correlation of PM and SARS-CoV2 infections

```{r, eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results = FALSE}

gam_lag_mean <- aggregate(gam_lag$t ~ gam_lag$lag, FUN = mean)
gam_lag_sd <- aggregate(gam_lag$t ~ gam_lag$lag, FUN = sd)
colnames(gam_lag_mean) <- c("lag", "t_mean")
colnames(gam_lag_sd) <- c("lag", "t_sd")

gam_lag_mean_sd <- merge(gam_lag_mean, gam_lag_sd)
gam_lag_mean_sd$sig <- NA
gam_lag_mean_sd$sig[match(tstat[tstat$p < 0.1, "lag"], gam_lag_mean_sd$lag)] <-
  gam_lag_mean_sd$t_mean[match(tstat[tstat$p < 0.1, "lag"], gam_lag_mean_sd$lag)]

figure_effectsize_errorbars <- ggplot(data = gam_lag_mean_sd, aes(x = as.factor(lag), y = t_mean)) +
  geom_point() +
  geom_point(aes(x = as.factor(lag), y = sig), color = "red") +
  geom_line() +
  geom_errorbar(aes(ymin = t_mean - t_sd, ymax = t_mean + t_sd)) +
  theme_bw() +
  geom_hline(yintercept = c(0), linetype = "dashed", color = "black") +
  labs(x = "Time lag [days]", y = "Effect size") +
  theme(
    text = element_text(size = 10), axis.title = element_text(size = 10),
    panel.background = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
figure_effectsize_errorbars

gam_lag_geo <- compileLaggedGLMGeo(gam_lag, cmpldata$de_nuts3_map)

ggplot(data = gam_lag_geo$test_geo[gam_lag_geo$test_geo$lag == -2, ]) +
  geom_point(aes(x = centroid_lat, y = t, color = as.factor(lag))) +
  theme_bw() +
  geom_hline(yintercept = c(-2, 2), linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(47, 55, 1)) +
  scale_y_continuous(breaks = seq(-10, 20, 2)) +
  labs(
    title = paste("Effect size of lagged ", pm),
    x = "Latitude [degrees]", y = "Effect size", color = "Lag of PM"
  )
```
